### 2020-10-25 outputs/paper/paper.Rmd
added first set of shibboleths

---
title: "Prediction of the 2020 Presidential Elections"
subtitle: "Spoiler alert: Biden Wins"
author: "Anees Shaikh, Jaffa Romain, and Lu Mu"
thanks: "Code and data are available at: https://github.com/aneesshake/elections_prediction"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: |
  | First sentence. Second sentence. Third sentence. Fourth sentence.
  |
  | **Keywords:** forecasting; US 2020 election; Trump;Biden
output:
  bookdown::pdf_document2:
    citation_package: natbib
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
```

# Introduction

# Data

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# Appendix {-}

\newpage

# References

### 2020-10-25 outputs/paper/paper.Rmd
Added the reference for the ACS and nationscape data. changed the directory for the scripts and included our info

@@
   |
-  | **Keywords:** forecasting; US 2020 election; Trump;Biden
+  | **Keywords:** forecasting; US 2020 election; Trump; Biden
 output:
@@
 # References
+
+Tausanovitch, Chris and Lynn Vavreck. 2020. Democracy Fund + UCLA Nationscape, October 10-17, 2019 (version 20200814). Retrieved from [URL].
+
+Steven Ruggles, Sarah Flood, Ronald Goeken, Josiah Grover, Erin Meyer, Jose Pacas and Matthew Sobek.
+IPUMS USA: Version 10.0 [dataset].
+Minneapolis, MN: IPUMS, 2020.
+https://doi.org/10.18128/D010.V10.0

### 2020-10-25 scripts/01-data_cleaning-post-strat.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Prepare and clean the survey data downloaded from [...UPDATE ME!!!!!]
-# Author: Rohan Alexander and Sam Caetano [CHANGE THIS TO YOUR NAME!!!!]
-# Data: 22 October 2020
-# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
+# Purpose: Prepare and clean the survey data downloaded from IPUMS. We will be downloading the ACS data from the aforementioned site
+# Author: Anees Shaikh, Jaffa Romain, Lu Mu
+# Data: 02 November 2020
+# Contact: anees.shaikh@mail.utoronto.ca or jaffa.romain@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-# - Need to have downloaded the ACS data and saved it to inputs/data
+# - Need to have downloaded the ACS data and saved it to inputs/raw_data/ACS
 # - Don't forget to gitignore it!
 
 
@@ -13,7 +13,7 @@
 library(haven)
 library(tidyverse)
 # Read in the raw data. 
-raw_data <- read_dta("inputs/data/usa_00003.dta"
+raw_data <- read_dta("inputs/raw_data/ACS/usa_00003.dta"
                      )
 # Add the labels
 raw_data <- labelled::to_factor(raw_data)

### 2020-10-25 scripts/01-data_cleaning-survey.R

@@ -1,12 +1,11 @@
 #### Preamble ####
-# Purpose: Prepare and clean the survey data downloaded from [...UPDATE ME!!!!!]
-# Author: Rohan Alexander and Sam Caetano [CHANGE THIS TO YOUR NAME!!!!]
-# Data: 22 October 2020
-# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
+# Purpose: Prepare and clean the survey data downloaded from IPUMS. We will be downloading the ACS data from the aforementioned site
+# Author: Anees Shaikh, Jaffa Romain, Lu Mu
+# Data: 02 November 2020
+# Contact: anees.shaikh@mail.utoronto.ca or jaffa.romain@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-# - Need to have downloaded the data from X and save the folder that you're 
-# interested in to inputs/data 
+# - Need to have downloaded the Nationscape survey data and saved it to inputs/raw_data/Nationscape
 # - Don't forget to gitignore it!
 
 
@@ -14,7 +13,7 @@
 library(haven)
 library(tidyverse)
 # Read in the raw data (You might need to change this if you use a different dataset)
-raw_data <- read_dta("inputs/data/ns20200625/ns20200625.dta")
+raw_data <- read_dta("inputs/raw_data/Nationscape/ns20200625.dta")
 # Add the labels
 raw_data <- labelled::to_factor(raw_data)
 # Just keep some variables

### 2020-10-27 scripts/01-data_cleaning-post-strat.R
Added cell counts to post strat script.

@@ -13,14 +13,25 @@
 library(haven)
 library(tidyverse)
 # Read in the raw data. 
-raw_data <- read_dta("inputs/raw_data/ACS/usa_00003.dta"
-                     )
+raw_data <- read_dta("inputs/raw_data/ACS/usa_00001.dta")
+
+raw_data %>% filter(age >= 18) %>% group_by(age) %>% summarise(n = sum(perwt))
+
+raw_data %>% filter(age >= 18) %>% group_by(age) %>%   tally()
+
+raw_data %>% group_by(educ) %>% tally()
+?summarise(n = sum(perwt))
+
 # Add the labels
 raw_data <- labelled::to_factor(raw_data)
 
 # Just keep some variables that may be of interest (change 
 # this depending on your interests)
+
+
 names(raw_data)
+
+raw_data
 
 reduced_data <- 
   raw_data %>% 
@@ -35,12 +46,24 @@ reduced_data <-
          citizen,
          educd,
          labforce,
-         inctot)
+         hhincome)
 rm(raw_data)
          
 
+
 #### What's next? ####
 
+##let us now create our different kinds of stratification
+##all filtering of the data based on the variables is done prior to this step
 
 
-         
+reduced_data %>% 
+  filter(age >= 18)
+  group_by(age) %>% 
+  tally()
+
+reduced_data %>% group_by(region) %>% tally()
+
+
+reduced_data %>% select(region) %>% distinct()         
+

### 2020-10-27 scripts/01-data_cleaning-survey.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Prepare and clean the survey data downloaded from IPUMS. We will be downloading the ACS data from the aforementioned site
+# Purpose: Prepare and clean the survey data downloaded from Nationscape. We will be downloading the nationscape data from the aforementioned site
 # Author: Anees Shaikh, Jaffa Romain, Lu Mu
 # Data: 02 November 2020
 # Contact: anees.shaikh@mail.utoronto.ca or jaffa.romain@mail.utoronto.ca

### 2020-10-27 scripts/ps4.R
Add files via upload

@@ -1,0 +1,59 @@
+# Uncomment these (by deleting the #) if you need to install the packages
+
+library(broom) # Helps make the regression results tidier
+library(here) # Helps make file referencing easier.
+library(tidyverse) # Helps make programming with R easier
+library(haven)
+
+library(stringr)
+
+
+# Read in the raw data (You might need to change this if you use a different dataset)
+raw_data <- read_dta("C:/Users/lumuc/Desktop/ps4/ns20200625/ns20200625.dta")
+# Add the labels
+raw_data <- labelled::to_factor(raw_data)
+# Just keep some variables
+data <- 
+  raw_data %>% 
+  select(vote_2020,
+         employment,
+         gender,
+         race_ethnicity,
+         household_income,
+         state,
+         age)
+
+head(raw_data)
+
+# Look at some summary statistics to make sure the data seem reasonable
+summary(raw_data)
+
+data <-
+  data %>%
+    filter(vote_2020 != "I am not sure/don't know")
+
+
+data <-
+  data %>%
+  filter(vote_2020 != "I would not vote")
+
+data <-
+  data %>%
+  mutate(vote_2020 = ifelse(vote_2020 == "Donald Trump", 1, 0))
+
+data %>% 
+  summarise(raw_prop = sum(vote_2020) / nrow(reduced_data))  
+
+model <- lm(vote_2020 ~ gender + age + employment + race_ethnicity + household_income + state,
+            data = data
+)
+
+broom::tidy(model)
+
+ACS_data <- read_dta("C:/Users/lumuc/Desktop/ps4/ns20200625/usa_00001.dta")
+head(ACS_data)
+
+ACS_data$estimate <-
+  model %>%
+  predict(newdata = ACS_data)
+

### 2020-10-28 scripts/01-data_cleaning-survey.R
added to model + survey cleaning

@@ -13,18 +13,20 @@
 library(haven)
 library(tidyverse)
 library(ggplot2)
-library(state)
+library(maps)
+library(ggthemes)
+library(mapproj)
 
-
-us_states <- map_data("state")
+us <- map_data("state")
+us<- us %>% mutate(state = case_when(state == "district of columbia"~"washington"))
 party_colors <- c("#FF4040","#0000FF")
 # Read in the raw data (You might need to change this if you use a different dataset)
-raw_data <- read_dta("inputs/raw_data/Nationscape/ns20200625.dta")
+survey_data<- read_dta("inputs/raw_data/Nationscape/ns20200625.dta")
 # Add the labels
-raw_data <- labelled::to_factor(raw_data)
+survey_data <- labelled::to_factor(survey_data)
 # Just keep some variables
-reduced_data <- 
-  raw_data %>% 
+survey_data_reduced <- 
+  survey_data %>% 
   select(interest,
          registration,
          vote_2016,
@@ -44,29 +46,37 @@ reduced_data <-
          age)
 
 # Keep observations of those who intend to vote in the upcoming 2020 election
-reduced_data <- reduced_data %>% filter(vote_intention == "Yes, I will vote")
+survey_data_reduced <-survey_data_reduced %>% filter(vote_intention == "Yes, I will vote")
 
 # Remove observations of those who are unsure of who they will vote for in 2020
-reduced_data <- reduced_data %>% filter(vote_2020 != "I am not sure/don't know")
+survey_data_reduced <-survey_data_reduced %>% filter(vote_2020 != "I am not sure/don't know")
 
 # We are predicting a binary outcome for vote; either Trump or Biden to win majority vote.
 # Keep vote results of either Democrat vs. Republican nominee
-reduced_data <- reduced_data %>% filter(vote_2020 == "Donald Trump"| vote_2020 == "Joe Biden")
+survey_data_reduced <-survey_data_reduced %>% filter(vote_2020 == "Donald Trump"| vote_2020 == "Joe Biden")
 
-# Age distribution
-reduced_data %>% ggplot(aes(x = age, fill  = vote_2020), color = "black") + geom_histogram(alpha = 0.7, color = "black", bins = 25) + 
[…]

### 2020-10-28 scripts/03-modelling.R

@@ -1,0 +1,42 @@
+# Uncomment these (by deleting the #) if you need to install the packages
+#### Preamble ####
+# Purpose: Prepare predictive model using individual survey data We will be downloading the ACS data from the aforementioned site
+# Author: Anees Shaikh, Jaffa Romain, Lu Mu
+# Data: 02 November 2020
+# Contact: anees.shaikh@mail.utoronto.ca or jaffa.romain@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# - Need to have downloaded the ACS data and saved it to inputs/raw_data/ACS
+# - Need to have loaded data in scripts folder; 01-data_cleaning_post_strat for post-stratification data and 01-data_cleaning-survey for survey data
+# in order to get cleaned data set.
+
+library(broom) # Helps make the regression results tidier
+library(here) # Helps make file referencing easier.
+library(tidyverse) # Helps make programming with R easier
+library(haven)
+
+library(stringr)
+
+# Look at some summary statistics to make sure the data seem reasonable
+summary(survey_data)
+
+# change vote_2020 results to binary result
+# 1: Trump, 0: Biden
+survey_data_reduced <-
+  survey_data_reduced %>%
+  mutate(vote_2020 = ifelse(vote_2020 == "Donald Trump", 1, 0))
+
+survey_data_reduced %>% 
+  summarise(raw_prop = sum(vote_2020) / nrow(survey_data_reduced))  
+
+model <- glm(vote_2020 ~ gender + age + employment + race_ethnicity + household_income + state,
+            data = survey_data_reduced)
+      
+
+broom::tidy(model)
+ACS_data <- reduced_data
+
+ACS_data$estimate <-
+  model %>%
+  predict(newdata = ACS_data)
+

### 2020-10-28 scripts/03-modelling.R
added to model

@@ -29,7 +29,7 @@ survey_data_reduced <-
 survey_data_reduced %>% 
   summarise(raw_prop = sum(vote_2020) / nrow(survey_data_reduced))  
 
-model <- glm(vote_2020 ~ gender + age + employment + race_ethnicity + household_income + state,
+model <- glm(vote_2020 ~ gender + age + employment + race_ethnicity + household_income + state, family=binomial,
             data = survey_data_reduced)
       
 

### 2020-10-30 outputs/paper/paper.Rmd
abstract commit

@@
 ---
-title: "Prediction of the 2020 Presidential Elections"
-subtitle: "Spoiler alert: Biden Wins"
-author: "Anees Shaikh, Jaffa Romain, and Lu Mu"
+title: "Prediction of the 2020 Presidential Election"
+subtitle: "Spoiler Alert: Biden Wins"
+author: "Anees Shaikh, Jaffa Romain, Lu Mu, and Cameron Fryer"
 thanks: "Code and data are available at: https://github.com/aneesshake/elections_prediction"
@@
 abstract: |
-  | First sentence. Second sentence. Third sentence. Fourth sentence.
+  | In this paper, we first consider the Democracy Fund + UCLA Nationscape Wave 50 dataset, which contains the results of a survey (conducted June 25-July 1, 2020) on American voter attitudes.
+Thereafter, the survey data is used to train a model relating voter intent to a few explanatory variables.
+The model is then applied to the post-stratification dataset; namely, the results of the 2018 1-year American Community Survey (ACS).
+Since the ACS data pertains to individual persons and their characteristics, the associated use of our model allows us to conclude that Joe Biden will be the next President of the United States.
   |
@@
 # Introduction
+
+To accurately predict the winner of the United States 2020 presidential election is difficult, as many different factors are involved.
+.....
 

### 2020-10-30 scripts/01-data_cleaning-survey.R
added to model + individual survey

@@ -18,7 +18,7 @@ library(ggthemes)
 library(mapproj)
 
 us <- map_data("state")
-us<- us %>% mutate(state = case_when(state == "district of columbia"~"washington"))
+us <- us %>% mutate(region = case_when(region == "district of columbia"~"washington", TRUE ~ region))
 party_colors <- c("#FF4040","#0000FF")
 # Read in the raw data (You might need to change this if you use a different dataset)
 survey_data<- read_dta("inputs/raw_data/Nationscape/ns20200625.dta")
@@ -43,7 +43,7 @@ survey_data_reduced <-
          education,
          state,
          congress_district,
-         age)
+         age, )
 
 # Keep observations of those who intend to vote in the upcoming 2020 election
 survey_data_reduced <-survey_data_reduced %>% filter(vote_intention == "Yes, I will vote")
@@ -55,16 +55,38 @@ survey_data_reduced <-survey_data_reduced %>% filter(vote_2020 != "I am not sure
 # Keep vote results of either Democrat vs. Republican nominee
 survey_data_reduced <-survey_data_reduced %>% filter(vote_2020 == "Donald Trump"| vote_2020 == "Joe Biden")
 
-# data with only democratic and republican nominees from 2016
-survey_2016 <- survey_data_reduced <-survey_data_reduced %>% filter(vote_2016 == "Donald Trump"| vote_2016 == "Hillary Clinton")
-  
-# Age distribution - 2016 elections
-survey_data_reduced %>% ggplot(aes(x = age, fill  = vote_2020), color = "black") + geom_histogram(alpha = 0.7, color = "black", bins = 25) + 
-  scale_fill_manual(values = c("#FF4040", "#0000FF")) + labs(title = "Age Distribution" , subtitle =  "Comparing age distribution of those who intend to vote for Trump and Biden.", x = "Age", y= "Number of votes", fill = "Nominee") + facet_wrap(vars(gender))
+# matching state names to that of the post-stratification data
+survey_data_reduced <- survey_data_reduced %>% mutate(state = case_when(state == "AL" ~ "alabama", state == "AK" ~ "alaska", state == "AZ" ~ "arizona", state == "AR" ~ "arkansas", state == "CA" ~ "california",
+                                                                        state == "CO" ~ "colorado", state == "CT" ~ "connecticut", state == "DE" ~ "delaware", state == "FL"~ "florida", state == "GA" ~ "georgia", 
[…]

### 2020-10-30 scripts/03-modelling.R

@@ -14,9 +14,7 @@ library(broom) # Helps make the regression results tidier
 library(here) # Helps make file referencing easier.
 library(tidyverse) # Helps make programming with R easier
 library(haven)
-
 library(stringr)
-
 # Look at some summary statistics to make sure the data seem reasonable
 summary(survey_data)
 
@@ -26,17 +24,31 @@ survey_data_reduced <-
   survey_data_reduced %>%
   mutate(vote_2020 = ifelse(vote_2020 == "Donald Trump", 1, 0))
 
+
 survey_data_reduced %>% 
-  summarise(raw_prop = sum(vote_2020) / nrow(survey_data_reduced))  
+  summarise(raw_prop = sum(vote_2020) / nrow(survey_data_reduced))  # no class bias for our response variable - results are fairy equal
 
-model <- glm(vote_2020 ~ gender + age + employment + race_ethnicity + household_income + state, family=binomial,
+# building model
+model <- glm(vote_2020 ~ sex + age + race + household_income + hispan + state, family=binomial(link = "logit"),
             data = survey_data_reduced)
-      
+summary(model)
 
+# Looking for indications of collinearity using correlation matrix of estimated coefficients
+car::vif(model) # no collinearity 
+# Influence diagnostic to check for any observations that might have an influence on the fitted model
 broom::tidy(model)
 ACS_data <- reduced_data
-
 ACS_data$estimate <-
   model %>%
   predict(newdata = ACS_data)
 
+
+### Testing Model
+set.seed(10)
+train <- survey_data_reduced %>% sample_frac(.70)
+predict <- predict(model,train, type = 'response')
+table_mat <- table(train$vote_2020, predict > 0.5)
+table_mat
+accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
+accuracy_Test
+# 63% - could be improved

### 2020-10-31 outputs/paper/paper.Rmd
added to model discussion + model finalized

@@
   |
-  | **Keywords:** forecasting; US 2020 election; Trump; Biden
+  | **Keywords:** forecasting; US 2020 election; Trump; Biden; multilevel regression with post-stratification
 output:
@@
 library(tidyverse)
-library(palmerpenguins)
+library(tinytex)
 ```
@@
 ```{r bills, fig.cap="Bills of penguins", echo = FALSE}
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = FALSE) +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+
 ```
@@
 
-# This needs to be about the random data tha tI sasmpled?
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = FALSE) +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
 ```
@@
 
-\begin{equation}
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
-\end{equation}
+In order to forecast the popular vote of the U.S 2020 elections, multilevel logistic regression with post-stratification (MRP) was used.
+The individual survey gives sample data on voters of the general U.S population, and post-stratification allows us to reweight estimates, adjusting bias between our sample and the target population, so that we have a representative sample of likely voters @ForecastingArticle.
+To achieve this, cells are constructed using variables in the cleaned Nationscape survey such as age, household income, and race.
+The model is then trained on the survey using the proportions given by the  ACS post-stratification data set.
+This approach gives us an advantage when attempted to forecast voting, as we can use a broad survey to speak to subsets in the population, and also tends to be less expensive to collect than non-probability samples @rohansmrp.
+However, using this approach places limitations on how much we can interpret from the estimates.
+Although we can estimate voting intention in different demographic groups, it doesn't tell us any qualitative information on voting patterns.
+For example, we might be able to see how different age categories vote, but we won't have a measure of the policy preferences, or their views on the candidates.
 
-Equation \@ref(eq:bayes) seems useful, eh?
+Using the Nationscape,  a multilevel logistic regression model can be fitted to the survey data set to model the proportion of voters who will vote for Trump.
+Logistic regression can be used to model a binary dependent variable, which is the choice between the Republic candidate, Donald Trump, and the Democratic candidate, Joe Biden in our case.
+This would be the main reason for choosing this model, over another model such as a linear regression model, where the output isn't necessarily binary.
+However, this imposes limitations.
+As we cannot take into consideration any other candidates running in the election due to the binary outcome.
+Another weakness is **** The variable 'vote_2020' in the data set was used as dependent variable, with the variable return 1 for voters intending to vote for Trump, and 0.
 
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
+Our model uses the variables pertaining to age, state, race, sex, household income, and whether a voter is Hispanic to predict our dependent variable. 
+The model takes the form:
+***  TO DO: Regression and Other Stories to cite equation **
+$$ Pr(y_i = 1) = logit^-1(X_i *\beta) $$
+where $y_i$ = 1 represents a voter voting for Trump in the election.
+$ Pr(y_i = 1) $ is the probability of a voter choosing to vote for Trump.
+The $ X_i\beta $ are the linear predictors, where $\beta$ represents the fitted coefficients for each independent variable $ X_i $ in the model.
+
+The logit function $logit(x) = log(x/1-x)$ maps the range (0, 1) to $(-\infinity, \infinity)$.
+Its inverse function $ logit^-1(x) = e^x/(1+e^x)$ maps back to the unit range. 
+The model's output is bounded between 0 and 1, unlike a linear regression model, mapping the outputs into a binary outcome, and says that the probability that some person $i$, will vote for Trump depends on their age, sex, state, household income,and whether the voter is Hispanic.
+
+We are running our regresion model using the 'glm()' function in [R]. The fitted model results in the following coefficients:
+```{r}
+
+```
 

### 2020-11-01 outputs/paper/paper.Rmd
re-wrote some code in survey cleaning to make code cleaner, and standardize outputs with post strat. Renamed post-strat script and standardized output added a few references and rough draft of discussion for paper section added 5th shibboleth

--- 2020-10-30 outputs/paper/paper.Rmd
@@
   |
-  | **Keywords:** forecasting; US 2020 election; Trump; Biden
+  | **Keywords:** forecasting; US 2020 election; Trump; Biden; Multilevel regression with post-stratification
 output:
@@
 
+All our analysis was done using R[@CiteR]
+
+Comparisons to ACS data and Nationscape data
+
+throw a few distributions in here to show the differences in distributions, and then tie each one in the discussion section to the potential implications of such a distribution.
+For instance, ACS data shows that there are more higher income earners than in Nationscape.
+Our model shows a strong effect between income and voting for a certain candidate.
+
+## comparison on age
+
+## comparison on race
+
+##
+
 # Discussion
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Talk about correlations between education and result
 
+To begin this discussion, we begin with a quick explanation of the US Federal voting system. This context is important for us to understand some key points and drawbacks with interpreting our model.
+
+Talk about previous research that has been done with similar models and the output of this model.
+Our model seems to indicate that Biden is going to win the popular vote.
+Throw some crap in here about 538 has certain polls that show a slight trump win.
+Show falling trump approvals and other things like that.
+This model has been thoroughly researched and has shown to be effective in multiple other contexts(needs citation)
+
+One of the key things we have to explain here is how the electoral college works.
+Readers may remember the 2016 US election where Hillary Clinton won the popular vote, however, this didn't mean that she won the federal election.
+This was due to her not winning the electoral college.
+The electoral college is a body that is elected to determine who will be the next president of the US.
+This happened in 2000 when Bush didn't win the popular vote, but did end up winning the electoral college
 ## Second discussion point

### 2020-11-01 scripts/01-data_cleaning-post-strat.R

@@ -13,14 +13,7 @@
 library(haven)
 library(tidyverse)
 # Read in the raw data. 
-raw_data <- read_dta("inputs/raw_data/ACS/usa_00001.dta")
-
-raw_data %>% filter(age >= 18) %>% group_by(age) %>% summarise(n = sum(perwt))
-
-raw_data %>% filter(age >= 18) %>% group_by(age) %>%   tally()
-
-raw_data %>% group_by(educ) %>% tally()
-?summarise(n = sum(perwt))
+raw_data <- read_dta("inputs/raw_data/ACS/usa_00001.dta") %>% filter(age >= 18)
 
 # Add the labels
 raw_data <- labelled::to_factor(raw_data)
@@ -57,13 +50,44 @@ rm(raw_data)
 ##all filtering of the data based on the variables is done prior to this step
 
 
-reduced_data %>% 
-  filter(age >= 18)
-  group_by(age) %>% 
+
+#counts by all vars
+reduced_data %>% group_by_all()
   tally()
 
-reduced_data %>% group_by(region) %>% tally()
+  
+#cell count by age
+reduced_data %>% 
+group_by(age) %>% 
+  tally()
 
+#cell count by region
+reduced_data %>% 
+  group_by(region) %>% 
+  tally()
 
-reduced_data %>% select(region) %>% distinct()         
+#cell count by sex
+reduced_data %>% 
+  group_by(sex) %>% 
+  tally()
+
+#cell count by race
+reduced_data %>% 
+  group_by(race) %>% 
+  tally()
+
+#cell count by education
+reduced_data %>% 
+  group_by(educd) %>% 
+  tally()
+
+#cell count by bpl
+reduced_data %>% 
+  group_by(bpl) %>% 
+  tally()
+
+#cell count by labforce
+reduced_data %>% 
+  group_by(labforce) %>% 
+  tally()
 

### 2020-11-01 scripts/01-data_cleaning-survey.R

@@ -20,10 +20,23 @@ library(mapproj)
 us <- map_data("state")
 us <- us %>% mutate(region = case_when(region == "district of columbia"~"washington", TRUE ~ region))
 party_colors <- c("#FF4040","#0000FF")
+
+
+#the little mapping below allows us to easily match state abbreviations to state names, thank you professor rohan alexander for suggesting this neat method.
+us_states_mapping <-  tibble(state.abb, state_name = tolower(state.name)) %>% 
+  rename(state = state.abb)
+
+
+
+                                   
+
+
 # Read in the raw data (You might need to change this if you use a different dataset)
 survey_data<- read_dta("inputs/raw_data/Nationscape/ns20200625.dta")
+
 # Add the labels
 survey_data <- labelled::to_factor(survey_data)
+
 # Just keep some variables
 survey_data_reduced <- 
   survey_data %>% 
@@ -43,76 +56,158 @@ survey_data_reduced <-
          education,
          state,
          congress_district,
-         age, )
+         age, ) %>% 
+  drop_na()
 
-# Keep observations of those who intend to vote in the upcoming 2020 election
-survey_data_reduced <-survey_data_reduced %>% filter(vote_intention == "Yes, I will vote")
 
-# Remove observations of those who are unsure of who they will vote for in 2020
-survey_data_reduced <-survey_data_reduced %>% filter(vote_2020 != "I am not sure/don't know")
 
-# We are predicting a binary outcome for vote; either Trump or Biden to win majority vote.
-# Keep vote results of either Democrat vs. Republican nominee
-survey_data_reduced <-survey_data_reduced %>% filter(vote_2020 == "Donald Trump"| vote_2020 == "Joe Biden")
+#creating this little ethnicity mapping to make the lower code cleaner
 
-# matching state names to that of the post-stratification data
-survey_data_reduced <- survey_data_reduced %>% mutate(state = case_when(state == "AL" ~ "alabama", state == "AK" ~ "alaska", state == "AZ" ~ "arizona", state == "AR" ~ "arkansas", state == "CA" ~ "california",
-                                                                        state == "CO" ~ "colorado", state == "CT" ~ "connecticut", state == "DE" ~ "delaware", state == "FL"~ "florida", state == "GA" ~ "georgia", 
[…]

### 2020-11-01 scripts/02-data_cleaning-post-strat.R

@@ -1,0 +1,180 @@
+#### Preamble ####
+# Purpose: Prepare and clean the survey data downloaded from IPUMS. We will be downloading the ACS data from the aforementioned site
+# Author: Anees Shaikh, Jaffa Romain, Lu Mu
+# Data: 02 November 2020
+# Contact: anees.shaikh@mail.utoronto.ca or jaffa.romain@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# - Need to have downloaded the ACS data and saved it to inputs/raw_data/ACS
+# - Don't forget to gitignore it!
+
+
+#### Workspace setup ####
+library(haven)
+library(tidyverse)
+
+
+
+
+
+#creating this list to determine whether an individual is a foreign or domestic born
+list_of_states <- tolower(state.name) %>% append("district of columbia")
+
+race_to_be_rewritten = c("black/african american/negro")
+
+races_to_be_filtered = c("three or more major races","two major races")
+
+
+
+# Read in the raw data. 
+raw_data <- 
+  read_dta("inputs/raw_data/ACS/usa_00001.dta") %>% 
+  filter(age >= 18 & hhincome <9999999 & hhincome >= 0) 
+  
+
+# Add the labels
+raw_data <- labelled::to_factor(raw_data) %>% filter(!citizen == "not a citizen")
+
+
+reduced_data <- 
+  raw_data %>% 
+  select(region,
+         stateicp,
+         sex, 
+         age, 
+         race, 
+         hispan, 
+         bpl,
+         citizen,
+         educd,
+         labforce,
+         hhincome) %>% drop_na()
+
+
+
+##let us now create our different kinds of stratification
+##all filtering of the data based on the variables is done prior to this step
+
+
+
+
+
+
+
+#determining whether a person is born in the US or outside.
+reduced_data <- 
+  reduced_data %>% 
+  mutate(foreign_born = if_else(bpl %in% list_of_states, "The United States", "Another country"))
+
+
+
+
+#removing respondents that have more than one race ascribed to them, as well as standardizing the black race name here
+reduced_data <- reduced_data %>% 
+  mutate(race = if_else(race %in% race_to_be_rewritten, "black", as.character(reduced_data$race))) %>% 
+  filter(!race %in% races_to_be_filtered)
+
+
+
+#grouping by age
+reduced_data<- reduced_data %>% 
+  mutate(age = case_when(as.numeric(age) <= 24 ~ "18-24", 
+                         as.numeric(age) <= 44 ~ "25-44",
[…]

### 2020-11-01 scripts/01-data_cleaning-post-strat.R
deleted old script name

@@ -1,93 +1,0 @@
-#### Preamble ####
-# Purpose: Prepare and clean the survey data downloaded from IPUMS. We will be downloading the ACS data from the aforementioned site
-# Author: Anees Shaikh, Jaffa Romain, Lu Mu
-# Data: 02 November 2020
-# Contact: anees.shaikh@mail.utoronto.ca or jaffa.romain@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: 
-# - Need to have downloaded the ACS data and saved it to inputs/raw_data/ACS
-# - Don't forget to gitignore it!
-
-
-#### Workspace setup ####
-library(haven)
-library(tidyverse)
-# Read in the raw data. 
-raw_data <- read_dta("inputs/raw_data/ACS/usa_00001.dta") %>% filter(age >= 18)
-
-# Add the labels
-raw_data <- labelled::to_factor(raw_data)
-
-# Just keep some variables that may be of interest (change 
-# this depending on your interests)
-
-
-names(raw_data)
-
-raw_data
-
-reduced_data <- 
-  raw_data %>% 
-  select(region,
-         stateicp,
-         sex, 
-         age, 
-         race, 
-         hispan,
-         marst, 
-         bpl,
-         citizen,
-         educd,
-         labforce,
-         hhincome)
-rm(raw_data)
-         
-
-
-#### What's next? ####
-
-##let us now create our different kinds of stratification
-##all filtering of the data based on the variables is done prior to this step
-
-
-
-#counts by all vars
-reduced_data %>% group_by_all()
-  tally()
-
-  
-#cell count by age
-reduced_data %>% 
-group_by(age) %>% 
-  tally()
-
-#cell count by region
-reduced_data %>% 
-  group_by(region) %>% 
-  tally()
-
-#cell count by sex
-reduced_data %>% 
-  group_by(sex) %>% 
-  tally()
-
-#cell count by race
-reduced_data %>% 
-  group_by(race) %>% 
-  tally()
-
-#cell count by education
-reduced_data %>% 
-  group_by(educd) %>% 
-  tally()
-
-#cell count by bpl
-reduced_data %>% 
-  group_by(bpl) %>% 
-  tally()
-
-#cell count by labforce
-reduced_data %>% 
-  group_by(labforce) %>% 
-  tally()
-

### 2020-11-01 outputs/paper/paper.Rmd
F

--- 2020-10-31 outputs/paper/paper.Rmd
@@
 abstract: |
-  | In this paper, we first consider the Democracy Fund + UCLA Nationscape Wave 50 dataset, which contains the results of a survey (conducted June 25-July 1, 2020) on American voter attitudes.
+  | The purpose of this analysis is to forecast the results of the upcoming United States 2020 presidential election.
+In this paper, we first consider the Democracy Fund + UCLA Nationscape Wave 50 dataset, which contains the results of a survey (conducted June 25-July 1, 2020) on American voter attitudes.
 Thereafter, the survey data is used to train a model relating voter intent to a few explanatory variables.
 The model is then applied to the post-stratification dataset; namely, the results of the 2018 1-year American Community Survey (ACS).
-Since the ACS data pertains to individual persons and their characteristics, the associated use of our model allows us to conclude that Joe Biden will be the next President of the United States.
-  |
-  | **Keywords:** forecasting; US 2020 election; Trump; Biden; multilevel regression with post-stratification
+Since the ACS data pertains to individual persons and their characteristics, the associated use of our model allows us to conclude that Joe Biden will be the next President of the United States.|
+Keywords: forecasting; US 2020 election; Trump; Biden; multilevel regression with post-stratification
 output:
@@
 library(tinytex)
+library(readr)
 ```
@@
 
-Our model uses the variables pertaining to age, state, race, sex, household income, and whether a voter is Hispanic to predict our dependent variable. 
+Our model uses the variables pertaining to age, state, race, sex, household income, and whether a voter is Hispanic to predict whether a respondent will vote for Trump or Biden in the elections. 
 The model takes the form:
@@
 $$ Pr(y_i = 1) = logit^-1(X_i *\beta) $$
-where $y_i$ = 1 represents a voter voting for Trump in the election.
-$ Pr(y_i = 1) $ is the probability of a voter choosing to vote for Trump.
-The $ X_i\beta $ are the linear predictors, where $\beta$ represents the fitted coefficients for each independent variable $ X_i $ in the model.
 
-The logit function $logit(x) = log(x/1-x)$ maps the range (0, 1) to $(-\infinity, \infinity)$.
-Its inverse function $ logit^-1(x) = e^x/(1+e^x)$ maps back to the unit range. 
-The model's output is bounded between 0 and 1, unlike a linear regression model, mapping the outputs into a binary outcome, and says that the probability that some person $i$, will vote for Trump depends on their age, sex, state, household income,and whether the voter is Hispanic.
+where $y_i = 1$ represents a voter voting for Trump in the election.
+$Pr(y_i = 1)$ is the probability of a voter choosing to vote for Trump.
+The $X_i\beta$ are the linear predictors, where $\beta$ represents the fitted coefficients for each independent variable $X_i$ in the model.
+The coefficient values signify the mean variable changes in a one unit shift in the given variable $\beta$.
+The logit function $logit(x) = log(x/1-x)$ maps the range (0, 1) to $(-\infty, \infty)$.
+Its inverse function $logit^-1(x) = e^x/(1+e^x)$ maps back to the unit range. 
+The model's output is bounded between 0 and 1,  mapping the outputs into a binary outcome.
+The output tells us the probability that some person $i$, will vote for Trump depends on their age, sex, state, household income, and whether the voter is Hispanic.
 
-We are running our regresion model using the 'glm()' function in [R]. The fitted model results in the following coefficients:
+We are running our regression model using the 'glm()' function in [R]. The fitted model results in the following coefficients:
 ```{r}
-
 ```
@@
 # Results
+
+All our analysis was done using R[@CiteR]
+
+Comparisons to ACS data and Nationscape data
+
+throw a few distributions in here to show the differences in distributions, and then tie each one in the discussion section to the potential implications of such a distribution.
+For instance, ACS data shows that there are more higher income earners than in Nationscape.
+Our model shows a strong effect between income and voting for a certain candidate.
+
+## comparison on age
+
+## comparison on race
+
+##
 

### 2020-11-01 scripts/03-modelling.R

@@ -1,7 +1,7 @@
 # Uncomment these (by deleting the #) if you need to install the packages
 #### Preamble ####
 # Purpose: Prepare predictive model using individual survey data We will be downloading the ACS data from the aforementioned site
-# Author: Anees Shaikh, Jaffa Romain, Lu Mu
+# Author: Anees Shaikh, Jaffa Romain, Lu Mu, Cameron Fryer
 # Data: 02 November 2020
 # Contact: anees.shaikh@mail.utoronto.ca or jaffa.romain@mail.utoronto.ca
 # License: MIT
@@ -25,25 +25,23 @@ survey_data_reduced <-
   survey_data_reduced %>%
   mutate(vote_2020 = ifelse(vote_2020 == "Donald Trump", 1, 0))
 
-
+saveRDS(survey_data_reduced, file = "inputs/survey_data.Rda")
 survey_data_reduced %>% 
   summarise(raw_prop = sum(vote_2020) / nrow(survey_data_reduced))  # no class bias for our response variable - results are fairy equal
 
 # building model
 model <- glm(vote_2020 ~ sex + age + race + household_income + hispan + state, family=binomial,
             data = survey_data_reduced)
-final_model <- saveRDS(model, file = "outputs/model/final_model.rds")
+saveRDS(model, file = "outputs/model/final_model.rds")
 # Looking for indications of collinearity using correlation matrix of estimated coefficients
 collin <- car::vif(model) # no collinearity 
-cor_test <- saveRDS(model, file = "outputs/model/cor.rds")
 
 
-coefficients <- broom::tidy(model)
-coef <- saveRDS(model, file = "outputs/model/coefficients.rds")
+coeffi <- broom::tidy(model) 
 
 
 
-### TTraining on indivual survey
+### Training on indivual survey
 set.seed(10)
 train <- survey_data_reduced %>% sample_frac(.70)
 predict <- predict(model,train, type = 'response')

### 2020-11-01 outputs/paper/paper.Rmd
added more references, saving rds locations for individual data and poststrat

@@
 library(readr)
+library(here)
 ```
@@
 ```{r}
+
+#Loading in the final model and the coefficents of the model in this cell.
+Please note that the outputs of those models are being stored in the outputs/model directory.
+The here package should give you the same result but if it doesn't then you may have to manually enter the same directory.
+final_model <- readRDS(here::here("outputs/model","final_model.rds"))
+coefficients <- readRDS(here::here("outputs/model","coefficients.rds"))
+
+cor <- readRDS(here::here("outputs/model","cor.rds"))
+
 ```

### 2020-11-01 scripts/01-data_cleaning-survey.R

@@ -107,6 +107,8 @@ survey_data_reduced <- survey_data_reduced %>%
                          age <= 64 ~ "45-64",  
                          age >= 65 ~ "65+")) 
 
+saveRDS(survey_data_reduced, file = "inputs/cleaned_data/individual-survey.rds")
+
 ### GRAPHING
 s <- survey_data_reduced %>% 
   select(vote_2016, race) %>% 

### 2020-11-01 scripts/02-data_cleaning-post-strat.R

@@ -38,17 +38,15 @@ raw_data <- labelled::to_factor(raw_data) %>% filter(!citizen == "not a citizen"
 
 reduced_data <- 
   raw_data %>% 
-  select(region,
+  select(
          stateicp,
          sex, 
          age, 
          race, 
          hispan, 
          bpl,
-         citizen,
-         educd,
-         labforce,
-         hhincome) %>% drop_na()
+         hhincome) %>% 
+  drop_na()
 
 
 
@@ -122,7 +120,7 @@ reduced_data <- reduced_data %>%
 
 
 
-
+saveRDS(reduced_data, file = "inputs/cleaned_data/post-strat.rds")
 
 
 
@@ -140,9 +138,9 @@ group_by(age) %>%
   summarise(num_records = n()) %>% 
   mutate(proportion = num_records/sum(num_records))
 
-#cell count by region
+#cell count by state
 reduced_data %>% 
-  group_by(region) %>% 
+  group_by(stateicp) %>% 
   summarise(num_records = n()) %>% 
   mutate(proportion = num_records/sum(num_records))
 
@@ -160,11 +158,6 @@ reduced_data %>%
   summarise(num_records = n()) %>% 
   mutate(proportion = num_records/sum(num_records))
 
-#cell count by education
-reduced_data %>% 
-  group_by(educd) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records))
 
 #cell count by bpl
 reduced_data %>% 
@@ -172,9 +165,4 @@ reduced_data %>%
   summarise(num_records = n()) %>% 
   mutate(proportion = num_records/sum(num_records))
 
-#cell count by labforce
-reduced_data %>% 
-  group_by(labforce) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records))
 

### 2020-11-01 scripts/02-data_cleaning-post-strat.R
added cell counts.rds to post-strat script

@@ -126,43 +126,55 @@ saveRDS(reduced_data, file = "inputs/cleaned_data/post-strat.rds")
 
 
 
+#counts by all vars
 
-hhincome #counts by all vars
-reduced_data %>% group_by_all()
-  tally()
+
+cell_counts <- reduced_data %>% 
+  group_by(stateicp,sex,age,race,hhincome_group,foreign_born) %>% 
+  summarise(num_records = n()) %>% ungroup() %>% 
+  mutate(proportion = num_records/sum(num_records),
+         total_num = sum(num_records)) %>% arrange(desc(proportion))
+
+
+saveRDS(cell_counts,file = "inputs/cleaned_data/cell_counts.rds")
 
   
 #cell count by age
 reduced_data %>% 
 group_by(age) %>% 
   summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records))
+  mutate(proportion = num_records/sum(num_records),
+         total_num = sum(num_records))
 
 #cell count by state
 reduced_data %>% 
   group_by(stateicp) %>% 
   summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records))
+  mutate(proportion = num_records/sum(num_records),
+         total_num = sum(num_records))
 
 
 #cell count by sex
 reduced_data %>% 
   group_by(sex) %>% 
   summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records))
+  mutate(proportion = num_records/sum(num_records),
+         total_num = sum(num_records))
 
 
 #cell count by race
 reduced_data %>% 
   group_by(race)%>% 
   summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records))
+  mutate(proportion = num_records/sum(num_records),
+         total_num = sum(num_records))
 
 
 #cell count by bpl
 reduced_data %>% 
   group_by(foreign_born) %>% 
   summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records))
+  mutate(proportion = num_records/sum(num_records),
+         total_num = sum(num_records))
 
 

### 2020-11-01 scripts/03-modelling.R

@@ -32,7 +32,12 @@ survey_data_reduced %>%
 # building model
 model <- glm(vote_2020 ~ sex + age + race + household_income + hispan + state, family=binomial,
             data = survey_data_reduced)
+
+
 saveRDS(model, file = "outputs/model/final_model.rds")
+
+
+
 # Looking for indications of collinearity using correlation matrix of estimated coefficients
 collin <- car::vif(model) # no collinearity 
 

### 2020-11-01 outputs/paper/paper.Rmd
added cell counts.rds to post-strat script
added hispanic to cell_counts

@@
 library(here)
+library(patchwork)
 ```
@@
 # Data
+
+```{r, echo=FALSE, include=FALSE}
+#loading in our 2 cleaned datasets
+
+individual_survey <- readRDS(here::here("inputs/cleaned_data","individual-survey.rds"))
+post_strat <- readRDS(here::here("inputs/cleaned_data","post-strat.rds"))
+
+```
 
@@
 
+(Figure \@ref(fig:age)) 
+```{r age, fig.cap="Difference in Age distribution between both datasets", echo = FALSE, fig.width=8, fig.height=4}
+
+individual_age_dist<- individual_survey %>% 
+  group_by(age) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "Nationscape")
+
+post_strat_age_dist <- post_strat %>% 
+  group_by(age) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "ACS")
+
+individual_age_dist %>% 
+  union(post_strat_age_dist) %>% 
+  ggplot(aes(x = age, y = percentage)) + 
+  geom_col() + 
+  facet_grid(~data_source) + 
+  theme_economist_white() +
+  labs(Title = )
+
+```
+
 ## comparison on race

### 2020-11-01 scripts/02-data_cleaning-post-strat.R

@@ -130,7 +130,7 @@ saveRDS(reduced_data, file = "inputs/cleaned_data/post-strat.rds")
 
 
 cell_counts <- reduced_data %>% 
-  group_by(stateicp,sex,age,race,hhincome_group,foreign_born) %>% 
+  group_by(stateicp,sex,age,race,hhincome_group,foreign_born,hispan) %>% 
   summarise(num_records = n()) %>% ungroup() %>% 
   mutate(proportion = num_records/sum(num_records),
          total_num = sum(num_records)) %>% arrange(desc(proportion))

### 2020-11-02 scripts/02-data_cleaning-post-strat.R
added to model + started predictions

@@ -66,7 +66,6 @@ reduced_data <-
 
 
 
-
 #removing respondents that have more than one race ascribed to them, as well as standardizing the black race name here
 reduced_data <- reduced_data %>% 
   mutate(race = if_else(race %in% race_to_be_rewritten, "black", as.character(reduced_data$race))) %>% 
@@ -87,7 +86,7 @@ reduced_data<- reduced_data %>%
 
 #income grouping, to match nationscape grouping
 reduced_data <- reduced_data %>%
-  mutate(hhincome_group = case_when(hhincome <= 14999 ~ "Less than $14,999",
+  mutate(household_income = case_when(hhincome <= 14999 ~ "Less than $14,999",
                                     hhincome <= 19999 ~ "$15,000 to $19,999",
                                     hhincome <= 24999 ~ "$20,000 to $24,999",
                                     hhincome <= 29999 ~ "$25,000 to $29,999",
@@ -118,8 +117,6 @@ reduced_data <- reduced_data %>%
 
 
 
-
-
 saveRDS(reduced_data, file = "inputs/cleaned_data/post-strat.rds")
 
 
@@ -130,11 +127,11 @@ saveRDS(reduced_data, file = "inputs/cleaned_data/post-strat.rds")
 
 
 cell_counts <- reduced_data %>% 
-  group_by(stateicp,sex,age,race,hhincome_group,foreign_born,hispan) %>% 
+  group_by(stateicp,sex,age,race,household_income, hispan) %>% 
   summarise(num_records = n()) %>% ungroup() %>% 
   mutate(proportion = num_records/sum(num_records),
          total_num = sum(num_records)) %>% arrange(desc(proportion))
-
+names(cell_counts)[names(cell_counts) == "stateicp"] <- "state_name"
 
 saveRDS(cell_counts,file = "inputs/cleaned_data/cell_counts.rds")
 

### 2020-11-02 scripts/03-modelling.R

@@ -16,6 +16,8 @@ library(tidyverse) # Helps make programming with R easier
 library(haven)
 library(stringr)
 library(brms)
+library(gtsummary)
+
 # Look at some summary statistics to make sure the data seem reasonable
 summary(survey_data)
 
@@ -30,7 +32,7 @@ survey_data_reduced %>%
   summarise(raw_prop = sum(vote_2020) / nrow(survey_data_reduced))  # no class bias for our response variable - results are fairy equal
 
 # building model
-model <- glm(vote_2020 ~ sex + age + race + household_income + hispan + state, family=binomial,
+model <- glm(vote_2020 ~ sex + age + race + household_income + hispan + state_name, family=binomial(link=logit),
             data = survey_data_reduced)
 
 
@@ -40,9 +42,9 @@ saveRDS(model, file = "outputs/model/final_model.rds")
 
 # Looking for indications of collinearity using correlation matrix of estimated coefficients
 collin <- car::vif(model) # no collinearity 
+c <- broom::tidy(model)
 
-
-coeffi <- broom::tidy(model) 
+saveRDS(c, file = "outputs/model/coefficient.rds")
 
 
 

### 2020-11-02 scripts/01-data_cleaning-survey.R
fixed mapping table

@@ -22,9 +22,14 @@ us <- us %>% mutate(region = case_when(region == "district of columbia"~"washing
 party_colors <- c("#FF4040","#0000FF")
 
 
-#the little mapping below allows us to easily match state abbreviations to state names, thank you professor rohan alexander for suggesting this neat method.
+
+
+
+#the little mapping below allows us to easily match state abbreviations to state names, thank you professor rohan alexander for suggesting this neat method.We're adding dc as a state here since both this survey and the ACS dataset regard it as such.
+#to do so, we're creating a small tibble that we'll union 
+district_columbia <- tibble(state = c("DC"), state_name =c("district of columbia"))
 us_states_mapping <-  tibble(state.abb, state_name = tolower(state.name)) %>% 
-  rename(state = state.abb)
+  rename(state = state.abb) %>% union(district_columbia)
 
 
 

### 2020-11-02 outputs/paper/paper.Rmd

--- 2020-11-01 outputs/paper/paper.Rmd
@@
 library(here)
+library(dplyr)
 ```
@@
 
-We are running our regression model using the 'glm()' function in [R]. The fitted model results in the following coefficients:
+We fit our regression model using the `glm()` function in [`R`].  We can see from table 1, which was made using `kable` from `knitr`,  that the fitted model results in the following coefficients:
 ```{r}
+# loading data sets and cell counts for MRP
+cell_counts <- readRDS(here::here("inputs/cleaned_data", "cell_counts.rds"))
+individual_survey <- readRDS(here::here("inputs/cleaned_data","individual-survey.rds"))
+post_strat <- readRDS(here::here("inputs/cleaned_data","post-strat.rds"))
 
@@
 final_model <- readRDS(here::here("outputs/model","final_model.rds"))
-coefficients <- readRDS(here::here("outputs/model","coefficients.rds"))
+coefs <- readRDS(here::here("outputs/model","coefficients.rds"))
+knitr::kable(c, caption = " Table 1: Model Coefficients") %>% kableExtra::kable_styling(latex_options = "scale_down")
+```
+In logistic regression, collinearity implies that we have predictor variables that are highly correlated; that is, they have a linear relationship.
+This is possible when you have a larger number of variables in a model, but having collinearity in multiple variables can lead to unstable estimates and inaccurate variances, which can affect confidence intervals.
+This could lead to incorrect inferences about the relationship between our response variable and explanatory variables @doi:10.1080/09720502.2010.10700699.
+Thus, it is imperative to check for any correlated variables in our model to avoid these issues.
+The value of the Variance Inflation Factor(VIF) in for the explanatory variables can give us a measure of collinearity our model.
+VIF values should be less than 5 to guarantee that collinearity is not an issue.
+We can see in table 2 below that all VIF values are less than 5, so collinearity will not be a concern for our model.
+```{r}
+cor <- readRDS(here::here("outputs/model","cor.rds"))
+knitr::kable(cor, caption =  "Checking Model Collinearity", digits = 2)
+```
 
-cor <- readRDS(here::here("outputs/model","cor.rds"))
+```{r}
+# make predictions using our model with data from post-strat data set
+
+cell_counts$estimate <- model %>% predict(newdata = cell_counts)
+
+cell_counts%>% mutate(predict_prop = estimate*proportion) %>% group_by(state_name) %>% summarize(alp_predict = sum(predict_prop))
 
@@
 
+*** add another diagnostic test
+
 # Results
+
+```{r, echo=FALSE, include=FALSE}
+
+```
 

### 2020-11-02 scripts/01-data_cleaning-survey.R
random spacing and merging jaffas changes

@@ -106,7 +106,7 @@ survey_data_reduced <- survey_data_reduced %>%
                          ifelse(hispanic == "Mexican", "mexican", 
                                 ifelse(hispanic == "Cuban", "cuban", 
                                        ifelse(hispanic == "Puerto Rican", "puerto rican", "other"))))) %>% 
-  mutate( sex = case_when(gender == "Female" ~ "female", gender == "Male"~ "male")) %>% 
+  mutate(sex = case_when(gender == "Female" ~ "female", gender == "Male"~ "male")) %>% 
   mutate(age = case_when(age <= 24 ~ "18-24", 
                          age <= 44 ~ "25-44",
                          age <= 64 ~ "45-64",  

### 2020-11-02 scripts/02-data_cleaning-post-strat.R

@@ -130,8 +130,10 @@ cell_counts <- reduced_data %>%
   group_by(stateicp,sex,age,race,household_income, hispan) %>% 
   summarise(num_records = n()) %>% ungroup() %>% 
   mutate(proportion = num_records/sum(num_records),
-         total_num = sum(num_records)) %>% arrange(desc(proportion))
-names(cell_counts)[names(cell_counts) == "stateicp"] <- "state_name"
+         total_num = sum(num_records)) %>% 
+  arrange(desc(proportion)) %>% 
+  rename(state_name = stateicp)
+
 
 saveRDS(cell_counts,file = "inputs/cleaned_data/cell_counts.rds")
 

### 2020-11-02 outputs/paper/paper.Rmd
RIP

@@
 library(dplyr)
+library(sjPlot)
 ```
@@
 The here package should give you the same result but if it doesn't then you may have to manually enter the same directory.
+
 final_model <- readRDS(here::here("outputs/model","final_model.rds"))
@@
 knitr::kable(c, caption = " Table 1: Model Coefficients") %>% kableExtra::kable_styling(latex_options = "scale_down")
+
 ```
@@
 ```{r}
+plot_model(final_model, vline.color = "red",  transform = NULL, title = "Figure _: Odds Ratios of the Logistic Regression model")
+```
+
+```{r}
 # make predictions using our model with data from post-strat data set
 
-cell_counts$estimate <- model %>% predict(newdata = cell_counts)
+cell_counts$estimate <- predict(final_model, newdata = cell_counts)
 
-cell_counts%>% mutate(predict_prop = estimate*proportion) %>% group_by(state_name) %>% summarize(alp_predict = sum(predict_prop))
+# post-stratified estimates - number of trump votes
+estimate_votes <- cell_counts %>% mutate(predict_prop = proportion * estimate) %>% group_by(state_name) %>% summarise(pred_vote = sum(predict_prop))
+estimate_votes
 
@@
 ```{r, echo=FALSE, include=FALSE}
-
+# by state
+estimates <- estimates %>% mutate(num_votes = n*estimate)
 ```

### 2020-11-02 outputs/paper/paper.Rmd
added a few graphs

--- 2020-11-02 outputs/paper/paper.Rmd
@@
 
-```{r, echo=FALSE, include=FALSE}
+```{r age, echo=FALSE, include=FALSE, fig.cap="Distribution of Age between ACS and Nationscape data"}
 
+individual_age_dist<- individual_survey %>% 
+  group_by(age) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "Nationscape")
+
+post_strat_age_dist <- post_strat %>% 
+  group_by(age) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "ACS")
+
+individual_age_dist %>% 
+  union(post_strat_age_dist) %>% 
+  ggplot(aes(x = age, y = percentage)) + 
+  geom_col() + 
+  facet_grid(~data_source) + 
+  theme_economist_white()
+```
+Looking at (Figure \@ref(fig:age)), we observe a shift in age distributions between the two datasets.
+There is a slight peak in the 25-44 age group in the nationscape data.
+We see this flatten and then observe a significant increase in the 65+ age group, and a minor increase in the 18-24 age groups.
+
+```{r}
+individual_hhincome_dist<- individual_survey %>% 
+  group_by(household_income) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "Nationscape") %>% rename(hhincome_group = household_income)
+
+post_strat_hhincome_dist <- post_strat %>% 
+  group_by(hhincome_group) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "ACS")
+
+individual_hhincome_dist %>% 
+  union(post_strat_hhincome_dist) %>% 
+  ggplot(aes(x = forcatshhincome_group, y = percentage)) + 
+  geom_col() + 
+  facet_grid(~data_source) + 
+  theme_economist_white()
 ```

### 2020-11-02 outputs/paper/paper.Rmd

--- 2020-11-02 outputs/paper/paper.Rmd
@@
 
+```{r age, echo=FALSE, include=FALSE, fig.cap="Distribution of Age between ACS and Nationscape data"}
+
+individual_age_dist<- individual_survey %>% 
+  group_by(age) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "Nationscape")
+
+post_strat_age_dist <- post_strat %>% 
+  group_by(age) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "ACS")
+
+individual_age_dist %>% 
+  union(post_strat_age_dist) %>% 
+  ggplot(aes(x = age, y = percentage)) + 
+  geom_col() + 
+  facet_grid(~data_source) + 
+  theme_economist_white()
+```
+Looking at (Figure \@ref(fig:age)), we observe a shift in age distributions between the two datasets.
+There is a slight peak in the 25-44 age group in the nationscape data.
+We see this flatten and then observe a significant increase in the 65+ age group, and a minor increase in the 18-24 age groups.
+
+```{r}
+individual_hhincome_dist<- individual_survey %>% 
+  group_by(household_income) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "Nationscape") %>% rename(hhincome_group = household_income)
+
+post_strat_hhincome_dist <- post_strat %>% 
+  group_by(hhincome_group) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "ACS")
+
+individual_hhincome_dist %>% 
+  union(post_strat_hhincome_dist) %>% 
+  ggplot(aes(x = forcatshhincome_group, y = percentage)) + 
+  geom_col() + 
+  facet_grid(~data_source) + 
+  theme_economist_white()
+
+```
+
 ```{r, echo=FALSE, include=FALSE}

### 2020-11-02 scripts/03-modelling.R
bunch of updates

@@ -18,22 +18,27 @@ library(stringr)
 library(brms)
 library(gtsummary)
 
+# loading in individual_data
+
+survey_data <- readRDS(here::here("inputs/cleaned_data","individual-survey.rds"))
+
 # Look at some summary statistics to make sure the data seem reasonable
+
+
 summary(survey_data)
 
 # change vote_2020 results to binary result
 # 1: Trump, 0: Biden
-survey_data_reduced <-
-  survey_data_reduced %>%
+survey_data <-
+  survey_data %>%
   mutate(vote_2020 = ifelse(vote_2020 == "Donald Trump", 1, 0))
 
-saveRDS(survey_data_reduced, file = "inputs/survey_data.Rda")
-survey_data_reduced %>% 
-  summarise(raw_prop = sum(vote_2020) / nrow(survey_data_reduced))  # no class bias for our response variable - results are fairy equal
+survey_data %>% 
+  summarise(raw_prop = sum(vote_2020) / nrow(survey_data))  # no class bias for our response variable - results are fairy equal
 
 # building model
 model <- glm(vote_2020 ~ sex + age + race + household_income + hispan + state_name, family=binomial(link=logit),
-            data = survey_data_reduced)
+            data = survey_data)
 
 
 saveRDS(model, file = "outputs/model/final_model.rds")
@@ -44,13 +49,13 @@ saveRDS(model, file = "outputs/model/final_model.rds")
 collin <- car::vif(model) # no collinearity 
 c <- broom::tidy(model)
 
-saveRDS(c, file = "outputs/model/coefficient.rds")
+saveRDS(c, file = "outputs/model/coefficients.rds")
 
 
 
 ### Training on indivual survey
 set.seed(10)
-train <- survey_data_reduced %>% sample_frac(.70)
+train <- survey_data %>% sample_frac(.70)
 predict <- predict(model,train, type = 'response')
 table_mat <- table(train$vote_2020, predict > 0.5)
 table_mat

### 2020-11-02 outputs/paper/paper.Rmd
idk
bunch of updates
push

@@
 # post-stratified estimates - number of trump votes
-estimate_votes <- cell_counts %>% mutate(predict_prop = proportion * estimate) %>% group_by(state_name) %>% summarise(pred_vote = sum(predict_prop))
+estimate_votes <- cell_counts %>% mutate(predict_prop = proportion * estimate) %>% group_by(state_name) %>% summarise(pred_vote = sum(predict_prop),
+                                                                                                                      mean_vote = mean(predict_prop))
 estimate_votes
+
+cell_counts
 
@@
 # Results
+
+```{r, echo=FALSE, include=FALSE}
+# by state
+estimates <- estimates %>% mutate(num_votes = n*estimate)
+```
 
@@
 
-```
-
-```{r, echo=FALSE, include=FALSE}
-# by state
-estimates <- estimates %>% mutate(num_votes = n*estimate)
 ```

### 2020-11-02 outputs/paper/paper.Rmd
fixing response type in the model

--- 2020-11-02 outputs/paper/paper.Rmd
@@
 coefs <- readRDS(here::here("outputs/model","coefficients.rds"))
-knitr::kable(c, caption = " Table 1: Model Coefficients") %>% kableExtra::kable_styling(latex_options = "scale_down")
+knitr::kable(coefs, caption = " Table 1: Model Coefficients") %>% kableExtra::kable_styling(latex_options = "scale_down")
 
@@
 
-cell_counts$estimate <- predict(final_model, newdata = cell_counts)
+cell_counts$estimate <- predict(final_model, newdata = cell_counts, type = "response")
 
 # post-stratified estimates - number of trump votes
-estimate_votes <- cell_counts %>% mutate(predict_prop = proportion * estimate) %>% group_by(state_name) %>% summarise(pred_vote = sum(predict_prop))
+estimate_votes <- cell_counts %>% 
+  mutate(predict_prop = proportion * estimate) %>% 
+  group_by(total_num,state_name) %>% 
+  summarise(predicted_vote = sum(predict_prop)) %>% 
+  mutate(state_prediction = total_num*predicted_vote)
+
 estimate_votes
+
+cell_counts %>% arrange(num_records)
 

### 2020-11-02 scripts/03-modelling.R

@@ -33,6 +33,10 @@ survey_data <-
   survey_data %>%
   mutate(vote_2020 = ifelse(vote_2020 == "Donald Trump", 1, 0))
 
+
+
+
+
 survey_data %>% 
   summarise(raw_prop = sum(vote_2020) / nrow(survey_data))  # no class bias for our response variable - results are fairy equal
 

### 2020-11-02 outputs/paper/paper.Rmd
gg

@@
 
+cell_counts
+
 cell_counts$estimate <- predict(final_model, newdata = cell_counts, type = "response")
@@
 estimate_votes <- cell_counts %>% 
-  mutate(predict_prop = proportion * estimate) %>% 
-  group_by(total_num,state_name) %>% 
-  summarise(predicted_vote = sum(predict_prop)) %>% 
-  mutate(state_prediction = total_num*predicted_vote)
+  mutate(predict_prop = prop * estimate) %>% 
+  group_by(state_name) %>% 
+  summarise(predicted_vote = sum(predict_prop)) %>% ungroup()
 
 estimate_votes
-
-cell_counts %>% arrange(num_records)
 

### 2020-11-02 scripts/02-data_cleaning-post-strat.R

@@ -124,56 +124,18 @@ saveRDS(reduced_data, file = "inputs/cleaned_data/post-strat.rds")
 
 
 #counts by all vars
-
-
 cell_counts <- reduced_data %>% 
-  group_by(stateicp,sex,age,race,household_income, hispan) %>% 
-  summarise(num_records = n()) %>% ungroup() %>% 
-  mutate(proportion = num_records/sum(num_records),
-         total_num = sum(num_records)) %>% 
-  arrange(desc(proportion)) %>% 
-  rename(state_name = stateicp)
+  group_by(stateicp, sex,age,race,hispan) %>% 
+  summarise(num_records = n()) %>%  
+  group_by(stateicp) %>% 
+  mutate(prop = num_records/sum(num_records),state_count = sum(num_records)) %>% ungroup() %>% rename(state_name = stateicp)
 
 
 saveRDS(cell_counts,file = "inputs/cleaned_data/cell_counts.rds")
 
-  
-#cell count by age
-reduced_data %>% 
-group_by(age) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records),
-         total_num = sum(num_records))
-
-#cell count by state
-reduced_data %>% 
-  group_by(stateicp) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records),
-         total_num = sum(num_records))
+ 
 
 
-#cell count by sex
-reduced_data %>% 
-  group_by(sex) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records),
-         total_num = sum(num_records))
 
-
-#cell count by race
-reduced_data %>% 
-  group_by(race)%>% 
-  summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records),
-         total_num = sum(num_records))
-
-
-#cell count by bpl
-reduced_data %>% 
-  group_by(foreign_born) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(proportion = num_records/sum(num_records),
-         total_num = sum(num_records))
 
 

### 2020-11-02 scripts/03-modelling.R

@@ -41,7 +41,7 @@ survey_data %>%
   summarise(raw_prop = sum(vote_2020) / nrow(survey_data))  # no class bias for our response variable - results are fairy equal
 
 # building model
-model <- glm(vote_2020 ~ sex + age + race + household_income + hispan + state_name, family=binomial(link=logit),
+model <- glm(vote_2020 ~ sex + age + race + hispan, family=binomial(link=logit),
             data = survey_data)
 
 

### 2020-11-02 outputs/paper/paper.Rmd
w or l?
first data commit (checking length)

@@
 
-Our data is of penguins (Figure \@ref(fig:bills)).
+As previously mentioned, our data is from two different datasets; the Democracy Fund + UCLA Nationscape Wave 50 dataset, and the 2018 1-year American Community Survey (ACS) dataset.
+They each contain both qualitative and quantitative data.
+In regard to the Nationscape Wave 50 dataset, a few of its strengths are the variety in variables considered, and the quality of the survey responses.
+With respect to its weaknesses, it is likely that the data contains sampling error and/or coverage error.
+One of the various strengths of the  ACS dataset is its ability to be customized with pre-specified (or pre-selected?) variables and sample size before download.
+Although there is a wide range of variables to choose from, being able to select only those which we needed made the exploratory data analysis process much more manageable than it could have otherwise been (with all possible variables).
+Furthermore, since sample size can be pre-specified to a limit of about 3.2 million people, the larger the sample size is set to, the greater the odds that the target population is represented with the data.
+In terms of weaknesses of the ACS dataset, there are none all that significant.
+Some minor concerns, however, are potential measurement error, processing error, and/or adjustment error in the data.
+  The Democracy Fund + UCLA Nationscape is "one of the largest public opinion surveys ever conducted -- interviewing people in nearly every county, congressional district, and mid-sized U.S.
+city in the leadup to the 2020 election." Consequently, interviewing "roughly 6,250 per week" was part of the survey methodology employed by Nationscape.
+As such, the Nationscape Wave 50 dataset contains results for 6,479 interviews conducted during the week of June 25-July 1, 2020.
+It can be accessed by visiting https://www.voterstudygroup.org/publication/nationscape-data-set and submitting a request to access the data.
+By providing your full name and email address, you will then be emailed a link to the page where it can be downloaded.
+  At first glance of the Wave 50 data, we noted that it contains 265 variables.
+This is noteworthy because it allows a more complete picture of voter attitude to be derived from the data compared to if there were less variables.
+Another strength of this dataset is the overall quality of responses from those surveyed.
+More specifically, the data does not contain any duplicate records, inappropriate entries, or poor entries (e.g.
+typos, misspellings); hence, making it easier to work with.
+This high quality of survey responses can be partially attributed to the fact that "Nationscape samples are provided by Lucid, a market research platform that runs an online exchange for survey respondents." What is more, the individuals belonging to the sample are required to "complete an attention check before completing the [online] survey." With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
+  The main weakness of the Nationscape Wave 50 dataset is the likelihood that it includes sampling error and/or coverage error.
+While the goal of the Nationscape survey is to determine American voter attitudes, making inferences based only on the Wave 50 dataset would result in sample bias.
+This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match "a set of
+demographic criteria." It is very unlikely that a sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
+This may lead one to wonder why Nationscape chose to use the described sampling approach.
+Apart from the fact that the survey was conducted every week, the reason is trade-offs.
+For example, although there is potential coverage error in the data--a common result of online surveys--some trade-offs are a lesser rate of non-response, lower cost, convenience and accuracy.
+Whilst on the topic of non-response, it is worth mentioning that the variable with greatest amount of non-response in the Wave 50 dataset is household income at rouhly 5.5%.
+Nonetheless, this is a much smaller percentage of non-response in household income than is typically seen in political surveys, and Nationscape handles it by not weighting income for non-respondents.
+Similarly, a statement from Nationscape in its "Representativeness Assessment" implies an overall lack of concern for non-response in the data, as "previous evaluations of the samples Lucid provides have found them to be of high quality."
+  The last aspect of the Nationscape Wave 50 dataset to be considered is the survey questionnaire itself.
+Strengths of the questionnaire include its structure, the design and wording of the questions being asked, and the variety of choices for closed-ended questions.
+As to the structure of the questionnaire, it contains an extensive amount of questions from which differing political views can be revealed.
+Furthermore, the survey questions are asked in an order that does not influence the answer of respondents to subsequent questions.
+In terms of the design and wording of the questions, they are closed-ended, and avoid the use of tricky wording.
+This is a strength of the questionnaire because it makes comparing the responses of individuals and thus, statistical analysis, much simpler than it would be with open-ended questions and/or troublesome language.
+The diverse number of options available to select as answers to the questions is a strength of the questionnaire because they are both mutually exclusive and exhaustive; hence, ensuring survey respondents are able to answer every question.
+With respect to weaknesses of the questionnaire, it seems that there are none other than its sheer length.
 
@@
 estimate_votes
+
+estimate_votes %>% mutate(status = if_else(predicted_vote>=0.50,"win","loss")) %>% group_by(status) %>% tally(
+)
 

### 2020-11-02 outputs/paper/paper.Rmd
added to model + plots for results and discussion

@@
 library(dplyr)
-library(sjPlot)
+library(ggplot2)
+library(ggthemes)
+library(ResourceSelection)
+library(maps)
+library(mapproj)
+library(caret)
 ```
@@
 
-To accurately predict the winner of the United States 2020 presidential election is difficult, as many different factors are involved.
 .....
@@
 
+ ## Individual Survey Data
+
+ ## Post-stratification Data
 As previously mentioned, our data is from two different datasets; the Democracy Fund + UCLA Nationscape Wave 50 dataset, and the 2018 1-year American Community Survey (ACS) dataset.
 They each contain both qualitative and quantitative data.
+To forecast the elections, we are using the Nationscape Wave 50 dataset.
 In regard to the Nationscape Wave 50 dataset, a few of its strengths are the variety in variables considered, and the quality of the survey responses.
@@
 
-```{r bills, fig.cap="Bills of penguins", echo = FALSE}
-
-```
-
-Talk more about it.
-
-Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)
-
-```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}
-
-```
-
-Talk way more about it.
-
 # Model
@@
 In order to forecast the popular vote of the U.S 2020 elections, multilevel logistic regression with post-stratification (MRP) was used.
-The individual survey gives sample data on voters of the general U.S population, and post-stratification allows us to reweight estimates, adjusting bias between our sample and the target population, so that we have a representative sample of likely voters @ForecastingArticle.
+MRP is useful for when generalizing from a possibly non-representative poll kennedy2020know.
+The individual survey gives sample data on voters of the general U.S population, and post-stratification allows us to re-weight estimates, adjusting bias between our sample and the target population, so that we have a representative sample of likely voters @ForecastingArticle.
 To achieve this, cells are constructed using variables in the cleaned Nationscape survey such as age, household income, and race.
@@
 
-Using the Nationscape,  a multilevel logistic regression model can be fitted to the survey data set to model the proportion of voters who will vote for Trump.
+Using the Nationscape data, a multilevel logistic regression model can be fitted to the survey data set to model the proportion of voters who will vote for Trump.
 Logistic regression can be used to model a binary dependent variable, which is the choice between the Republic candidate, Donald Trump, and the Democratic candidate, Joe Biden in our case.
@@
 As we cannot take into consideration any other candidates running in the election due to the binary outcome.
-Another weakness is **** The variable 'vote_2020' in the data set was used as dependent variable, with the variable return 1 for voters intending to vote for Trump, and 0.
+The variable 'vote_2020' in the data set was used as the dependent variable (the variable we are trying to predict), with the variable returning 1 for voters intending to vote for Trump, and 0 if they are voting for Biden.
 
@@
 We fit our regression model using the `glm()` function in [`R`].  We can see from table 1, which was made using `kable` from `knitr`,  that the fitted model results in the following coefficients:
-```{r}
+```{r, echo=FALSE}
 # loading data sets and cell counts for MRP
@@
 coefs <- readRDS(here::here("outputs/model","coefficients.rds"))
-knitr::kable(coefs, caption = " Table 1: Model Coefficients") %>% kableExtra::kable_styling(latex_options = "scale_down")
+knitr::kable(coefs, caption = " Table 1: Model Coefficients", digits = 2)
 
@@
 We can see in table 2 below that all VIF values are less than 5, so collinearity will not be a concern for our model.
-```{r}
+```{r, echo=FALSE}
 cor <- readRDS(here::here("outputs/model","cor.rds"))
@@
 
-```{r}
-plot_model(final_model, vline.color = "red",  transform = NULL, title = "Figure _: Odds Ratios of the Logistic Regression model")
+```{r, echo=FALSE}
+ htst <- hoslem.test(final_model$y, fitted(final_model), g=10)
+htst
 ```
+Before the model can be used for any predictions, we must also check that our model fits our data well, and meets any assumptions made by the model.
+The Hosmer-Lemeshow goodness of fit test assesses whether or not the observed events match expected events in subgroups of the model.
+For this test, the p-value can be used to assess goodness of fit.
+A smaller p-value indicates strong evidence against a good fit.
+In the case of our model, we can see in table 4 that we have a very large p-value of 0.9, indicating no evidence of poor fit.
 
-```{r}
+| Table 4: Hosmer and Lemeshow goodness of fit (GOF) test 	|    	|         	|
+|---------------------------------------------------------	|----	|---------	|
+| X-squared                                               	| df 	| p-value 	|
+| 3.5                                                     	| 8  	| 0.9     	|
+
+```{r, echo=FALSE}
 # make predictions using our model with data from post-strat data set
@@
 cell_counts$estimate <- predict(final_model, newdata = cell_counts, type = "response")
+#  prediction intervals
+# fitted value - error to get margin of error
+se_md <- predict(final_model, newdata = cell_counts, type = "response", se.fit = T)
 
-# post-stratified estimates - number of trump votes
-estimate_votes <- cell_counts %>% 
-  mutate(predict_prop = prop * estimate) %>% 
+ lwr <-  cell_counts$estimate - 1.96*se_md$se.fit
+ upper <- cell_counts$estimate + 1.96*se_md$se.fit
+
+# all estimates with lower and upper bounds
+estimates_with_margins <- cell_counts %>% mutate(n = num_records * estimate) %>% cbind(cell_counts, lwr, upper)
+# removing duplicate column name estimate
+estimates_with_margins <- estimates_with_margins[, !duplicated(colnames(estimates_with_margins))]
+
+# post-stratified estimates - number of trump votes with prediction intervals
+post_strat_estimates <- estimates_with_margins %>% 
+  mutate(predict_prop = prop * estimate, l_predict = lwr * prop, u_predict = upper * prop) %>% 
   group_by(state_name) %>% 
-  summarise(predicted_vote = sum(predict_prop)) %>% ungroup()
+  summarise(predicted_vote = sum(predict_prop), lower = sum(l_predict), upp = sum(u_predict)) %>% group_by(state_name) %>% ungroup()
 
-estimate_votes
-
-estimate_votes %>% mutate(status = if_else(predicted_vote>=0.50,"win","loss")) %>% group_by(status) %>% tally(
+post_strat_estimates %>% mutate(status = if_else(predicted_vote>=0.50,"win","loss")) %>% group_by(status) %>% tally(
 )
@@
 
-*** add another diagnostic test
-
 # Results
 
-```{r, echo=FALSE, include=FALSE}
-# by state
-estimates <- estimates %>% mutate(num_votes = n*estimate)
-```
-
 ```{r age, echo=FALSE, include=FALSE, fig.cap="Distribution of Age between ACS and Nationscape data"}
-
 individual_age_dist<- individual_survey %>% 
@@
 
-```{r}
-individual_hhincome_dist<- individual_survey %>% 
-  group_by(household_income) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(percentage = 100 * num_records/sum(num_records),
-         total_num = sum(num_records),
-         data_source = "Nationscape") %>% rename(hhincome_group = household_income)
+```{r, echo = FALSE}
 
-post_strat_hhincome_dist <- post_strat %>% 
-  group_by(hhincome_group) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(percentage = 100 * num_records/sum(num_records),
-         total_num = sum(num_records),
-         data_source = "ACS")
+# estimated proportion of votes by State
+post_strat_estimates %>%
+ggplot(aes(y = predicted_vote, x =state_name, color = "MRP estimate")) + geom_point() +
+geom_errorbar(aes(ymin = lower, ymax = upp), width = 0) +
+ylab("Proportion of Trump Supporters") +
+xlab("State") +
+geom_point(data = individual_survey %>%
+group_by(state_name, vote_2020) %>% summarise(n = n()) %>% group_by(state_name) %>%
+mutate(prop = n/sum(n)) %>% filter(vote_2020==1),
+aes(state_name, prop, color = "Raw data")) + theme_minimal() +
+scale_color_brewer(palette = "Set1") + theme(legend.position = "bottom") + theme(legend.title = element_blank()) + 
+  theme(axis.text.x = element_text(angle = 35, hjust = 1, size = 5)) + coord_flip() +labs(title = "2020 Predicted Popular Vote by State", subtitle = "Compating redicted popular vote by state using raw data and MRP")
+```
 
-individual_hhincome_dist %>% 
-  union(post_strat_hhincome_dist) %>% 
-  ggplot(aes(x = forcatshhincome_group, y = percentage)) + 
-  geom_col() + 
-  facet_grid(~data_source) + 
-  theme_economist_white()
+```{r, echo = FALSE}
+# Sorting State wins
+election_wins <- post_strat_estimates %>% mutate(winner = if_else(predicted_vote>=0.50,"Trump","Biden")) %>% group_by(state_name
+)
+election_wins$state <- str_to_title(as.character(election_wins$state_name))
+election_wins <- election_wins %>% filter(state != "district of columbia")
+
+# Plotting state wins
+statebins(election_wins, value_col="winner", ggplot2_scale_function = scale_fill_manual, values = c(Trump = "#2166ac", Biden = "#b2182b"), name = "state") +
+  theme_statebins(legend_position="right")
 

### 2020-11-02 scripts/01-data_cleaning-survey.R

@@ -111,6 +111,11 @@ survey_data_reduced <- survey_data_reduced %>%
                          age <= 44 ~ "25-44",
                          age <= 64 ~ "45-64",  
                          age >= 65 ~ "65+")) 
+survey_data_reduced <-
+  survey_data_reduced %>%
+  mutate(vote_2020 = ifelse(vote_2020 == "Donald Trump", 1, 0))
+
+
 
 saveRDS(survey_data_reduced, file = "inputs/cleaned_data/individual-survey.rds")
 

### 2020-11-02 scripts/03-modelling.R

@@ -29,10 +29,6 @@ summary(survey_data)
 
 # change vote_2020 results to binary result
 # 1: Trump, 0: Biden
-survey_data <-
-  survey_data %>%
-  mutate(vote_2020 = ifelse(vote_2020 == "Donald Trump", 1, 0))
-
 
 
 
@@ -54,17 +50,5 @@ collin <- car::vif(model) # no collinearity
 c <- broom::tidy(model)
 
 saveRDS(c, file = "outputs/model/coefficients.rds")
-
-
-
-### Training on indivual survey
-set.seed(10)
-train <- survey_data %>% sample_frac(.70)
-predict <- predict(model,train, type = 'response')
-table_mat <- table(train$vote_2020, predict > 0.5)
-table_mat
-accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
-accuracy_Test
-# 63% - could be improved
-
+  
 

### 2020-11-02 outputs/paper/paper.Rmd
added comments to each file. added refs

@@
 library(caret)
+library(statebins)
 ```
@@
  ## Post-stratification Data
-As previously mentioned, our data is from two different datasets; the Democracy Fund + UCLA Nationscape Wave 50 dataset, and the 2018 1-year American Community Survey (ACS) dataset.
+As previously mentioned, our data is from two different datasets; the Democracy Fund + UCLA Nationscape Wave 50 dataset(Tausanovitch and Vavreck, 2020), and the 2018 1-year American Community Survey (ACS) dataset(Ruggles et al, 2020).
 They each contain both qualitative and quantitative data.
@@
 This is possible when you have a larger number of variables in a model, but having collinearity in multiple variables can lead to unstable estimates and inaccurate variances, which can affect confidence intervals.
-This could lead to incorrect inferences about the relationship between our response variable and explanatory variables @doi:10.1080/09720502.2010.10700699.
+This could lead to incorrect inferences about the relationship between our response variable and explanatory variables [@doi:10.1080/09720502.2010.10700699].
 Thus, it is imperative to check for any correlated variables in our model to avoid these issues.
@@
 
-post_strat_estimates %>% mutate(status = if_else(predicted_vote>=0.50,"win","loss")) %>% group_by(status) %>% tally(
-)
+post_strat_estimates <- post_strat_estimates %>% 
+  mutate(biden_vote_proportion = 1-predicted_vote,
+         conservative_election_status = if_else(lower %>% between(0.48,0.52),"toss-up",
+                                                 if_else(lower > 0.52, "win","loss")),
+         liberal_election_status  = if_else(upp %>% between(0.48,0.52),"toss-up",
+                                                 if_else(upp > 0.52, "win","loss")),
+         middle_ground_election_status = if_else(predicted_vote %>% between(0.48,0.52),"toss-up",
+                                                 if_else(predicted_vote > 0.52, "win","loss")))
 
@@
 
+All our analysis was done using R[@CiteR].
+
+The graphs for our variables are done in similar fashion for each of our variables.
+First, the difference in their distributions between both datasets is displayed.
+Then, showing the relationship between those variables and voting preferences is shown.
+The purpose of such a display is to showcase the effect of the variables, and then how a change in the distribution will affect the result.
+For instance, we may observe that women are more likely to vote for Biden, and then subsequently we notice that there are more women in the ACS census.
+We will go in-depth about the implications of these observations in the discussion section.
+
+## comparison on age
 ```{r age, echo=FALSE, include=FALSE, fig.cap="Distribution of Age between ACS and Nationscape data"}
@@
 
+## comparison on race
+
+```{r}
+individual_sex_dist<- individual_survey %>% 
+  group_by(sex) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "Nationscape")
+
+post_strat_sex_dist <- post_strat %>% 
+  group_by(sex) %>% 
+  summarise(num_records = n()) %>% 
+  mutate(percentage = 100 * num_records/sum(num_records),
+         total_num = sum(num_records),
+         data_source = "ACS")
+
+individual_sex_dist %>% 
+  union(post_strat_sex_dist) %>% 
+  ggplot(aes(x = sex, y = percentage)) + 
+  geom_col() + 
+  facet_grid(~data_source) + 
+  theme_economist_white()
+
+```
+
+```{r}
+#figure to see 2020 voting by sex
+vote_2020_sex <- individual_survey %>% 
+  mutate(vote_2020 = if_else(vote_2020 == 1,"Donald Trump","Joe Biden")) %>% 
+  group_by(sex,vote_2020) %>% 
+  tally() %>% 
+  ggplot(aes(x=sex, y = n)) + 
+  geom_col() + 
+  facet_grid(~vote_2020)
+
+#figure to see 2016 voting by sex
+vote_2016_sex <- individual_survey %>% 
+  group_by(sex,vote_2016) %>% 
+  tally() %>% 
+  ggplot(aes(x=sex, y = n)) + 
+  geom_col() + 
+  facet_grid(~vote_2016)
+
+library(patchwork)
+vote_2020_sex + vote_2016_sex
+```
+
+Figure (@fig) indicates that a man is much more likely to vote for Trump than a woman is. This tren
+
+(@fig)
 ```{r, echo = FALSE}
@@
 post_strat_estimates %>%
-ggplot(aes(y = predicted_vote, x =state_name, color = "MRP estimate")) + geom_point() +
+ggplot(aes(y = predicted_vote, x = state_name, color = "MRP estimate")) + geom_point() +
 geom_errorbar(aes(ymin = lower, ymax = upp), width = 0) +
@@
 scale_color_brewer(palette = "Set1") + theme(legend.position = "bottom") + theme(legend.title = element_blank()) + 
-  theme(axis.text.x = element_text(angle = 35, hjust = 1, size = 5)) + coord_flip() +labs(title = "2020 Predicted Popular Vote by State", subtitle = "Compating redicted popular vote by state using raw data and MRP")
+  theme(axis.text.x = element_text(angle = 35, hjust = 1, size = 5)) + coord_flip() +labs(title = "2020 Predicted Popular Vote by State", subtitle = "Comparing predicted popular vote by state using raw data and MRP")
 ```
 
-```{r, echo = FALSE}
+```{r heatmap, echo = FALSE, fig.cap="Heatmap showcasing model predictions at a state-level"}
 # Sorting State wins
-election_wins <- post_strat_estimates %>% mutate(winner = if_else(predicted_vote>=0.50,"Trump","Biden")) %>% group_by(state_name
-)
+election_wins <- 
+  post_strat_estimates %>% 
+  mutate(winner = if_else(predicted_vote>=0.50,"Trump","Biden")) %>% 
+  group_by(state_name)
+
 election_wins$state <- str_to_title(as.character(election_wins$state_name))
+
 election_wins <- election_wins %>% filter(state != "district of columbia")
@@
 # Plotting state wins
-statebins(election_wins, value_col="winner", ggplot2_scale_function = scale_fill_manual, values = c(Trump = "#2166ac", Biden = "#b2182b"), name = "state") +
+statebins(election_wins, value_col="winner", ggplot2_scale_function = scale_fill_manual, values = c(Trump ="#b2182b" , Biden = "#2166ac"), name = "state") +
   theme_statebins(legend_position="right")
@@
 
-All our analysis was done using R[@CiteR]
+Figure(\@ref(fig:heatmap)) shows the predictions of our model on the post-stratified dataset. This figure ignores close calls and has a binary Trump or Biden result.
 
-Comparisons to ACS data and Nationscape data
-
-throw a few distributions in here to show the differences in distributions, and then tie each one in the discussion section to the potential implications of such a distribution.
-For instance, ACS data shows that there are more higher income earners than in Nationscape.
-Our model shows a strong effect between income and voting for a certain candidate.
+```{r}
+post_strat_estimates %>% group_by(conservative_election_status) %>% summarise(num_states = n()) %>% rename(election_status = conservative_election_status) %>%  knitr::kable(caption = "Outcome of election")
+```
 

### 2020-11-02 scripts/01-data_cleaning-survey.R

@@ -17,7 +17,10 @@ library(maps)
 library(ggthemes)
 library(mapproj)
 
+
+##Static variables##
 us <- map_data("state")
+
 us <- us %>% mutate(region = case_when(region == "district of columbia"~"washington", TRUE ~ region))
 party_colors <- c("#FF4040","#0000FF")
 
@@ -66,8 +69,8 @@ survey_data_reduced <-
 
 
 
-#creating this little ethnicity mapping to make the lower code cleaner
-
+#creating this little ethnicity mapping to make the lower code cleaner, we'll join our survey_data_reduced later to make the ethnicities consistent with ACS data
+#as opposed to writing out every level, we're extracting our levels from the dataset.
 race_ethnicity <- attr(survey_data_reduced$race_ethnicity, "levels")
 us_ethnicity_mapping <- data.frame("race" = c("white",
                                               "black",
@@ -86,6 +89,7 @@ us_ethnicity_mapping <- data.frame("race" = c("white",
                                               "other race, nec"),race_ethnicity)
 
 
+#
 survey_data_reduced <-survey_data_reduced %>% 
   filter(vote_intention == "Yes, I will vote") %>% # Keep observations of those who intend to vote in the upcoming 2020 election
   filter(vote_2020 == "Donald Trump"| vote_2020 == "Joe Biden") %>% # Since we're only looking for a binary result, we're filtering for respondents that would've voted for Trump or Biden. By nature of the filter conditions, we'd be removing all other types of responses.
@@ -119,12 +123,17 @@ survey_data_reduced <-
 
 saveRDS(survey_data_reduced, file = "inputs/cleaned_data/individual-survey.rds")
 
-### GRAPHING
+### GRAPHING ###
+
+#these graph were done as EDA to see if there were any notable trends
 s <- survey_data_reduced %>% 
   select(vote_2016, race) %>% 
   group_by(race, vote_2016) %>% 
   summarise(n = n()) %>% 
   mutate(freq = n/sum(n))
+
+
+
 
 
 ## 2016 vote by race

### 2020-11-02 scripts/02-data_cleaning-post-strat.R

@@ -1,13 +1,13 @@
 #### Preamble ####
-# Purpose: Prepare and clean the survey data downloaded from IPUMS. We will be downloading the ACS data from the aforementioned site
-# Author: Anees Shaikh, Jaffa Romain, Lu Mu
-# Data: 02 November 2020
+# Purpose: Prepare and clean the data downloaded from IPUMS for our post-stratification cell counts. We will be downloading the ACS data from https://usa.ipums.org/usa-action/variables/group
+# Author: Anees Shaikh, Jaffa Romain, Lu Mu, and Cameron Fryer
+# Date: 02 November 2020
 # Contact: anees.shaikh@mail.utoronto.ca or jaffa.romain@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
 # - Need to have downloaded the ACS data and saved it to inputs/raw_data/ACS
-# - Don't forget to gitignore it!
-
+# - Don't forget to gitignore it and cite it!
+# - It would also be helpful for you to know how the values for variables differ between both datasets. This way you can lay out a simple plan to match them.
 
 #### Workspace setup ####
 library(haven)
@@ -16,26 +16,34 @@ library(tidyverse)
 
 
 
+## Defined static Variables ##
+#These variables may change according to your needs, you may find that you have no interest in the foreign born status, adjust the variables accordingly
 
 #creating this list to determine whether an individual is a foreign or domestic born
 list_of_states <- tolower(state.name) %>% append("district of columbia")
 
+#the black/african american race is defined differently in the nationscape dataset, so we're going to use this variable for transformation later
 race_to_be_rewritten = c("black/african american/negro")
 
+#For our analysis, we're not looking at people who may be defined by multiple races, this classification doesn't mesh well with the nationscape data
 races_to_be_filtered = c("three or more major races","two major races")
 
 
 
-# Read in the raw data. 
+# Read in the raw data. You may choose to use the here::here() function if it works better with your setup. Do remember that this won't work unless you have the file extracted into this folder.
+#Individuals below 18 can't vote and are filtered accordingly
+#according to the IPUMS codes, an income of "9999999" is used to signify a N/A value. we are also removing negative values for income as that didn't make sense to us. There didn't seem to be any information about it either in the codes
 raw_data <- 
   read_dta("inputs/raw_data/ACS/usa_00001.dta") %>% 
   filter(age >= 18 & hhincome <9999999 & hhincome >= 0) 
   
 
-# Add the labels
[…]

### 2020-11-02 scripts/03-modelling.R

@@ -1,6 +1,6 @@
 # Uncomment these (by deleting the #) if you need to install the packages
 #### Preamble ####
-# Purpose: Prepare predictive model using individual survey data We will be downloading the ACS data from the aforementioned site
+# Purpose: Prepare predictive model using individual survey data and ACS data. Both these datasets have been prepared and cleaned using 01-data_cleaning-survey.R and 02-data_cleaning-post-strat.R. You must run those beforehand
 # Author: Anees Shaikh, Jaffa Romain, Lu Mu, Cameron Fryer
 # Data: 02 November 2020
 # Contact: anees.shaikh@mail.utoronto.ca or jaffa.romain@mail.utoronto.ca
@@ -18,7 +18,7 @@ library(stringr)
 library(brms)
 library(gtsummary)
 
-# loading in individual_data
+# loading in cleaned individual data
 
 survey_data <- readRDS(here::here("inputs/cleaned_data","individual-survey.rds"))
 
@@ -36,7 +36,7 @@ summary(survey_data)
 survey_data %>% 
   summarise(raw_prop = sum(vote_2020) / nrow(survey_data))  # no class bias for our response variable - results are fairy equal
 
-# building model
+# building model, a logistic regression where vote_2020 is modelled by sex, age, race, hispan
 model <- glm(vote_2020 ~ sex + age + race + hispan, family=binomial(link=logit),
             data = survey_data)
 

### 2020-11-03 outputs/paper/paper.Rmd
model + data + references done

@@
 title: "Prediction of the 2020 Presidential Election"
-subtitle: "Spoiler Alert: Biden Wins"
+subtitle: "Spoiler Alert: Trump Wins"
 author: "Anees Shaikh, Jaffa Romain, Lu Mu, and Cameron Fryer"
@@
 In this paper, we first consider the Democracy Fund + UCLA Nationscape Wave 50 dataset, which contains the results of a survey (conducted June 25-July 1, 2020) on American voter attitudes.
-Thereafter, the survey data is used to train a model relating voter intent to a few explanatory variables.
-The model is then applied to the post-stratification dataset; namely, the results of the 2018 1-year American Community Survey (ACS).
-Since the ACS data pertains to individual persons and their characteristics, the associated use of our model allows us to conclude that Joe Biden will be the next President of the United States.|
+Thereafter, the survey data is used to train our model relating voter intent to 4 explanatory variables.
+The model is then applied to the post-stratification dataset; namely, the results of the 2018 1-year American Community Survey (ACS), to get predictions representative of the target population.
+Since the ACS data pertains to individual persons and their characteristics, the associated use of our model allows us to conclude that Donald Trump will be relected as the President of the United States, winning the popular vote by ___ with a margin of error of ______
 Keywords: forecasting; US 2020 election; Trump; Biden; multilevel regression with post-stratification
@@
 
- ## Post-stratification Data
-As previously mentioned, our data is from two different datasets; the Democracy Fund + UCLA Nationscape Wave 50 dataset(Tausanovitch and Vavreck, 2020), and the 2018 1-year American Community Survey (ACS) dataset(Ruggles et al, 2020).
+As previously mentioned, our data is from two different datasets; the Democracy Fund + UCLA Nationscape Wave 50 dataset, and the 2018 1-year American Community Survey (ACS) dataset.
 They each contain both qualitative and quantitative data.
-To forecast the elections, we are using the Nationscape Wave 50 dataset.
-In regard to the Nationscape Wave 50 dataset, a few of its strengths are the variety in variables considered, and the quality of the survey responses.
-With respect to its weaknesses, it is likely that the data contains sampling error and/or coverage error.
-One of the various strengths of the  ACS dataset is its ability to be customized with pre-specified (or pre-selected?) variables and sample size before download.
-Although there is a wide range of variables to choose from, being able to select only those which we needed made the exploratory data analysis process much more manageable than it could have otherwise been (with all possible variables).
-Furthermore, since sample size can be pre-specified to a limit of about 3.2 million people, the larger the sample size is set to, the greater the odds that the target population is represented with the data.
-In terms of weaknesses of the ACS dataset, there are none all that significant.
-Some minor concerns, however, are potential measurement error, processing error, and/or adjustment error in the data.
+To forecast the results of the presidential election, our model was fitted using the Nationscape survey data.
 The Democracy Fund + UCLA Nationscape is "one of the largest public opinion surveys ever conducted -- interviewing people in nearly every county, congressional district, and mid-sized U.S.
-city in the leadup to the 2020 election." Consequently, interviewing "roughly 6,250 per week" was part of the survey methodology employed by Nationscape.
+city in the lead up to the 2020 election." \cite{vavreck_tausanovitch_2020} Consequently, interviewing "roughly 6,250 per week" was part of the survey methodology employed by Nationscape.
 As such, the Nationscape Wave 50 dataset contains results for 6,479 interviews conducted during the week of June 25-July 1, 2020.
-It can be accessed by visiting https://www.voterstudygroup.org/publication/nationscape-data-set and submitting a request to access the data.
+It can be accessed by visiting https://www.voterstudygroup.org/publication/nationscape-data-set and submitting a request to access the data \cite{vavreck_tausanovitch_2020}.
 By providing your full name and email address, you will then be emailed a link to the page where it can be downloaded.
+
+Nationscape samples are provided by Lucid, a market research platform that uses their software for conducting online surveys.
+The Nationscape surveys were completed by participants online in English \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019} .
+The survey data is then weighted using a simple raking technique, where one set of weights is generated for each week's survey, in an attempt to have data that is representative of their target population.
+The survey data is weighted on many demographic factors such as gender, age, and nativity.
+
   At first glance of the Wave 50 data, we noted that it contains 265 variables.
@@
 typos, misspellings); hence, making it easier to work with.
-This high quality of survey responses can be partially attributed to the fact that "Nationscape samples are provided by Lucid, a market research platform that runs an online exchange for survey respondents." What is more, the individuals belonging to the sample are required to "complete an attention check before completing the [online] survey." With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
+This high quality of survey responses can be at least partially attributed to the fact that "Nationscape samples are provided by Lucid, a market research platform that runs an online exchange for survey respondents.
+\cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}" What is more, the individuals belonging to the sample are required to "complete an attention check before completing the [online] survey \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}." With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
+
   The main weakness of the Nationscape Wave 50 dataset is the likelihood that it includes sampling error and/or coverage error.
 While the goal of the Nationscape survey is to determine American voter attitudes, making inferences based only on the Wave 50 dataset would result in sample bias.
-This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match "a set of
-demographic criteria." It is very unlikely that a sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
+This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match "a set of demographic criteria \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}." It is unlikely that a survey sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
 This may lead one to wonder why Nationscape chose to use the described sampling approach.
@@
 For example, although there is potential coverage error in the data--a common result of online surveys--some trade-offs are a lesser rate of non-response, lower cost, convenience and accuracy.
-Whilst on the topic of non-response, it is worth mentioning that the variable with greatest amount of non-response in the Wave 50 dataset is household income at rouhly 5.5%.
+Whilst on the topic of non-response, it is worth mentioning that the variable with greatest amount of non-response in the Wave 50 data set is household income at roughly 5.5%.
 Nonetheless, this is a much smaller percentage of non-response in household income than is typically seen in political surveys, and Nationscape handles it by not weighting income for non-respondents.
-Similarly, a statement from Nationscape in its "Representativeness Assessment" implies an overall lack of concern for non-response in the data, as "previous evaluations of the samples Lucid provides have found them to be of high quality."
+Similarly, a statement from Nationscape in its "Representativeness Assessment" implies an overall lack of concern for non-response in the data, as "previous evaluations of the samples Lucid provides have found them to be of high quality \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}."
+
   The last aspect of the Nationscape Wave 50 dataset to be considered is the survey questionnaire itself.
 Strengths of the questionnaire include its structure, the design and wording of the questions being asked, and the variety of choices for closed-ended questions.
-As to the structure of the questionnaire, it contains an extensive amount of questions from which differing political views can be revealed.
+As to the structure of the questionnaire, it contains an extensive amount of questions from which differing political views are revealed -- the ideal circumstance for reflecting on American voter attitudes.
 Furthermore, the survey questions are asked in an order that does not influence the answer of respondents to subsequent questions.
@@
 The diverse number of options available to select as answers to the questions is a strength of the questionnaire because they are both mutually exclusive and exhaustive; hence, ensuring survey respondents are able to answer every question.
-With respect to weaknesses of the questionnaire, it seems that there are none other than its sheer length.
+With regard to weaknesses of the questionnaire, it seems that there are none other than its sheer length.
+
+ The data was cleaned to get our target population of eligible voters that intend to vote in the 2020 elections.
+Voters who didn't plan to vote or not eligible were filtered from the data set, and the variable pertaining to the 2020 election vote was cleaned to consist only of voters who plan on voting for either Trump or Biden.
+Before the process of fitting our model, we selected variables of interest that we felt gave insight on how different demographics vote.
+The resulting data set consisted of 3541 observations, and later reduced to the variables state, age, sex, Hispanic, race, and intended 2020 vote.
+
+```{r ageind, echo=FALSE, fig.cap="Distribution of Age in Nationscape data"}
+individual_survey <- readRDS(here::here("inputs/cleaned_data","individual-survey.rds"))
+post_strat <- readRDS(here::here("inputs/cleaned_data","post-strat.rds"))
+
+cont <- mutate(individual_survey, vote = (case_when(individual_survey$vote_2020 == 1~"Trump", individual_survey$vote_2020 == 0~"Biden")))
+
+  ggplot(cont, aes(x = age, fill  = vote), color = "black") + 
+  geom_bar(alpha = 0.7, color = "black") + 
+  scale_fill_manual(values = c("#0000FF","#FF4040")) + 
+  labs(title = "Age Vote Distribution" , 
+       subtitle =  " A look into how eligible voters intend to vote based on age.", 
+       x = "Age", 
+       y= "Number of votes", 
+       fill = "Nominee") +   
+  theme(axis.text.x = element_text(angle = 35, hjust = 1, size = 5))
+
+```
+
+Age:  The age of a respondent.
+Age was originally a nominal value(numerical), but the data was later grouped into categories for our predictive model, and to match the post-stratification data.
+This allowed for us to look at how age groups might influence the vote.
+The 2020 election marks the first time that Millennials and Gen Z will have an equal share of the number of eligible voters as Baby Boomers and earlier generations \cite{@agearticle} .
+Older voters in the past have had higher turn out rates, so looking at how how age categories tend to vote can give insight on how this will play a factor in the future election.
+We can see in (Figure \@ref(fig:ageind)) that there is a much smaller proportion of younger voters, but this is the age group that seems to have the strongest Democratic support.
+
+Sex: The sex of a respondent.
+Certain policies currently being such as those on anti-Abortion largely impact the female population in the United States.
+It seems that sex can be an important indicator of how voters respond to the platforms of the parties with regards to sex-specific policies.
+
+Hispanic: This variable looks at whether a respondent is Hispanic or not.
+In the 2020 election, Latinos will be the largest racial/ethnic minority in the electoral.
+With Trump's previous comments on countries such as Mexico, it seems important to see how pivotal the vote of this group will be in the election \cite{@citeBBC}.
+
+```{r racedis, echo=FALSE, fig.cap= "Votes by Race in Nationscape data"}
+#these graph were done as EDA to see if there were any notable trends
+s <-cont %>% 
+  select(vote_2016, race) %>% 
+  group_by(race, vote_2016) %>% 
+  summarise(n = n()) %>% 
+  mutate(freq = n/sum(n))
+## 2016 vote by race
+s %>% ggplot(aes(x = vote_2016, fill = race, group = race)) + 
+  geom_bar(aes(y = freq), stat = "identity", position = "dodge", alpha = 2/3) + 
+  labs(x = "Race", title = " Percentage of 2016 Presidential Election Votes by Race")
+```
+
+Race: (Figure \@ref(fig:racedis)) shows the distribution of votes in the 2016 federal election; specifically the Democratic candidate, Hillary Clinton, and the Republican candidate, Donald Trump.
+We can see that certain groups had a significant different in their voting preference.
+for example, majority of Black voters voted for Clinton rather than Trump, as well as the population of Chinese voters.
+
+State: When forecasting the election outcome, we will be predicting by State to see the majority vote in each.
+Certain states have been dominated by one particular party.
+For example, the Democratic party hasn't won the majority vote in the presidential election since 1992 \cite{{@cite270towin}.
+Since the election is based on electoral colleges as well, the number of Sates won by a candidate should give us a clear prospect for the dominating candidate.
+
+ ## Post-stratification Data
+
+  For our next step in our analysis, we use post-stratification to overcome any bias and have data representative of our target population.
+Conducted on a yearly basis, the American Community Survey "is a nationwide survey that collects and produces information on social, economic, housing, and demographic characteristics" of the United States population.
+In order to do so, the United States Census Bureau "mails questionnaires to approximately 295,000 addresses a month across the [country]." It should be noted that all the addresses receiving a letter to participate in the ACS are from a random sample chosen by the Census Bureau each month.
+Furthermore, every single address in the United States has approximately a 1 in 480 chance of being randomly selected to participate in the ACS during a given month.
+The exception to this comes from addresses that have already been selected to participate in the last 5 years; they are not allowed to be selected for another monthly sample.
+  To access the 2018 1-year ACS dataset, the first step is to visit https://usa.ipums.org/usa/ and create an account.
+From there, IPUMS will send you an email to verify your email address.
+By clicking the link contained in the email, you will be redirected to the IPUMS website and your email will be verified.
+Once on the website, you will be able to access the page where ACS sample and variables can be selected.
+To choose the 2018 1-year ACS as your sample, press "Select Samples".
+For this analysis, the single 2018 sample was chosen.
+The website allows for the user to choose the sample, as well as any variables of interest.
+After selecting the sample, you will be taken to a page wherein variable categories are listed.
+In addition, there is also an option to search for variables of interest.
+The variables that we selected for the dataset used in this paper are: region, stateicp, hhincome, sex, age, marst, race, hispan, bpl, citizen, educ, labforce, and ftotinc.
+When finished selecting variables, click "View Cart" near the top right of the page, and you will arrive at the cart page.
+This is your chance to look over all of the chosen variables.
+If you are happy with them, press the blue button near the middle of the page that reads "Create Data Extract." On the new page that you arrive at, there is a blue button near the bottom of the page with "Customize Sample Sizes" displayed in it.
+If you wish to change the size of your sample, then click the button.
+Now that our dataset has been fully customized, press the blue "Submit Extract" button at the very bottom of the page.
+Finally, you have reached the page from which your dataset can be downloaded, and the final step is to wait for the sample to be ready.
+
+  Regarding the survey methodology employed by the Census Bureau for the ACS, they do not find people to survey, but instead find households.
+The target population of the ACS is the population of the United States, and the sampling frame consists of all Americans living at a specific address.
+The sample itself includes all Americans living at one of the 295,000 addresses that are selected to participate in the ACS during a given month.
+Furthermore, the Census Bureau designs the sample in a way that guarantees favourable geographic inclusion, while also utilizing random sampling to select the "addresses to be included in the ACS." Every resident of the addresses which the invitation to complete the ACS was mailed is required to complete either the online survey, or "to mail the completed paper questionnaire." If neither requirement is satisfied after about a month, the Census Bureau "will mail an additional paper survey questionnaire." If the second paper questionnaire goes uncompleted too, then roughly 1 month after it was sent, someone will show up at your address to conduct an in-person interview with you.
+Due to the fact that completing the ACS is the law in the United States, non-response does not appear to be a very big issue.
+In terms of the ACS sampling approach, there are definitely some trade-offs made.
+For instance, although random sampling can cause sampling bias in the data, this concern is outweighed by the benefits of the sampling approach used.
+A few obvious benefits of the sampling approach that come to mind are a lower cost of administering the ACS, as well as a widespread geographic representation in the data.
+
+  In addition to the strength of the ACS dataset with respect to its customizability, another is the timeliness of the data.
+To be precise, survey data was continuously collected throughout 2018, and the results were shared in 2019.
+This is very important as it means there is opportunity to gain a more thorough understanding of current characteristics in the United States than has ever before been possible.
+One of the only obstacles in achieving this comes from the main weakness of the ACS dataset--poor data quality.
+For example, there are numerous duplicate records and inappropriate entries contained in the data.
+Some potential explanations for the low standard of data are measurement error, processing error, and/or adjustment error.
+More explicitly, measurement error is a possible reason for the bad data because when American citizens are selected to complete the ACS survey, they are required to complete it by law.
+Although the survey can be completed online, or the completed questionnaire can be mailed back, for someone who never wanted to complete the survey to begin with, it would not be a big surprise for them to quickly record poor quality answers and submit it.
+Processing error could be contained in the data as a result of either an error in coding with respect to the online survey, or an error in data entry by the person inputting completed paper questionnaires to the computer.
+Lastly, adjustment error serves as a possible explanation for the low quality data since any adjustments made to past survey results could have had an adverse affect.
 
@@
 In order to forecast the popular vote of the U.S 2020 elections, multilevel logistic regression with post-stratification (MRP) was used.
-MRP is useful for when generalizing from a possibly non-representative poll kennedy2020know.
-The individual survey gives sample data on voters of the general U.S population, and post-stratification allows us to re-weight estimates, adjusting bias between our sample and the target population, so that we have a representative sample of likely voters @ForecastingArticle.
+MRP is useful for when generalizing from a possibly non-representative poll \cite{kennedy2020know}.
+The individual survey gives sample data on voters of the general U.S population, and post-stratification allows us to re-weight estimates, adjusting bias between our sample and the target population, so that we have a representative sample of likely voters \cite{alexander2020a}.
 To achieve this, cells are constructed using variables in the cleaned Nationscape survey such as age, household income, and race.
 The model is then trained on the survey using the proportions given by the  ACS post-stratification data set.
-This approach gives us an advantage when attempted to forecast voting, as we can use a broad survey to speak to subsets in the population, and also tends to be less expensive to collect than non-probability samples @rohansmrp.
+This approach gives us an advantage when attempted to forecast voting, as we can use a broad survey to speak to subsets in the population, and also tends to be less expensive to collect than non-probability samples \cite{alexander2020a}.
 However, using this approach places limitations on how much we can interpret from the estimates.
@@
 The model takes the form:
-***  TO DO: Regression and Other Stories to cite equation **
 $$ Pr(y_i = 1) = logit^-1(X_i *\beta) $$
+Equation from: \cite{@RAOS}
 
@@
 The coefficient values signify the mean variable changes in a one unit shift in the given variable $\beta$.
+
 The logit function $logit(x) = log(x/1-x)$ maps the range (0, 1) to $(-\infty, \infty)$.
@@
 cell_counts <- readRDS(here::here("inputs/cleaned_data", "cell_counts.rds"))
-individual_survey <- readRDS(here::here("inputs/cleaned_data","individual-survey.rds"))
-post_strat <- readRDS(here::here("inputs/cleaned_data","post-strat.rds"))
 
@@
 This is possible when you have a larger number of variables in a model, but having collinearity in multiple variables can lead to unstable estimates and inaccurate variances, which can affect confidence intervals.
-This could lead to incorrect inferences about the relationship between our response variable and explanatory variables [@doi:10.1080/09720502.2010.10700699].
+This could lead to incorrect inferences about the relationship between our response variable and explanatory variables.
 Thus, it is imperative to check for any correlated variables in our model to avoid these issues.
@@
 
-All our analysis was done using R[@CiteR].
+All our analysis was done using R \cite{@CiteR}.
 

### 2020-11-03 outputs/paper/paper.Rmd
made lots of changes to paper, references

--- 2020-11-02 outputs/paper/paper.Rmd
@@
 Since the ACS data pertains to individual persons and their characteristics, the associated use of our model allows us to conclude that Joe Biden will be the next President of the United States.|
-Keywords: forecasting; US 2020 election; Trump; Biden; multilevel regression with post-stratification
+Keywords: Forecasting; US 2020 election; Trump; Biden; Multilevel regression with post-stratification
 output:
@@
 
-## comparison on race
+## comparison on sex
 
@@
 
-```{r}
+```{r sex_vote,fig.cap="Which individual did the candidate vote for? 2020 on the left, 2016 on the right"}
 #figure to see 2020 voting by sex
@@
   geom_col() + 
-  facet_grid(~vote_2020)
+  facet_grid(~vote_2020)+ labs(x = "Sex", y = "Number of people")
 
@@
   geom_col() + 
-  facet_grid(~vote_2016)
+  facet_grid(~vote_2016) + labs(x = "Sex", y = "Number of people")
 
@@
 
-Figure (@fig) indicates that a man is much more likely to vote for Trump than a woman is. This tren
+Figure (@fig) indicates that a man is much more likely to vote for Trump than a woman is. This trend is also observed in the 2016 elections when Hillary Clinton, a woman, ran for office.
 
-(@fig)
+```{r survey_election_results}
+individual_survey %>% mutate(vote_2020 = if_else(vote_2020 == 1,"Donald Trump","Joe Biden")) %>% group_by(vote_2020) %>% tally() %>% rename(candidate = vote_2020, votes = n ) %>% knitr::kable(caption = "Polling results for 2020 Election")
+```
+The above table indicates that Trump is most likely to win. This is captured from Nationscape data.
+
+```{r model2020prediction}
+post_strat_estimates %>% group_by(middle_ground_election_status) %>% tally() %>% rename(state_result = middle_ground_election_status, number = n) %>%  knitr::kable(caption = "Model Prediction for 2020")
+```
+
 ```{r, echo = FALSE}
@@
 
-```{r}
-post_strat_estimates %>% group_by(conservative_election_status) %>% summarise(num_states = n()) %>% rename(election_status = conservative_election_status) %>%  knitr::kable(caption = "Outcome of election")
-```
-
-## comparison on age
-
-## comparison on race
-
-##
-
 # Discussion
 
-## First discussion point
+While American citizens play a significant role in electing the next President, it is safe to say that the results of the US federal election are watched closely by the entire world.
+It goes without saying that the results of this election will have implications around the world.
+To contextualize the results of our model and analysis, we need a quick and simplified lesson on the American electoral process.
+Unlike most democracies, such as Canada, the head of state is typically elected by a popular vote.
+In the US, however, the head of state is not elected by a popular vote.
+Instead, citizens vote for a representative, or an elector, that then elects the president.
+This group of electors forms the electoral college.
+The electoral college is tasked with electing the next President.
+If a candidate wins a state by even a small percent, they effectively win all the electors in the state.
 
-Talk about correlations between education and result
+with that background knowledge in mind, what do our poll results indicate?
+From Figure(\@ref(fig:survey_election_results)), we gather that Trump has a very slight lead.
+Multiple other polls have shown Biden to be in the lead(FiveThirtyEight, 2020).
+Now there are a few reasons as to why we may not want to use poll results to predict elections.
+Polls have of late garnered a lot of criticism due to their incorrect predictions of the 2016 Federal election results[@citeAapor].
+Note that while they got the ultimate results wrong, they did get the popular vote prediction correct[@citeAapor].Nearly every national poll that year had Hillary Clinton winning, with some estimates of her having a 3% point lead[@citeAapor].
+Ultimately, she won the popular vote by 2.1%.
+However, as mentioned previously, the US does not use a popular vote system.
+At the state level, the polls became less effective, indicating a Clinton lead again[@citeAapor].
+State level polls do translate over to the electoral college a bit more effectively than national polls do.
+Another reason we may not want to overly rely on polls is yet another cautious tale from the 2016 election season.
+In that season, polls had over-represented college-graduates, a group that was more likely to vote for Clinton, thus tipping the scales even more in favour of Clinton[@citeAapor].
+Capturing this over-representation is key to effectively predicting an election.
+As mentioned in the results section, some results were intentionally shown as their distributions and then their voting preferences.
+This was done to elucidate the effect of a difference in representation, akin to  the difference of representation of educated voters between polled data and actual populations.
+From figure(\@ref(fig:age)), there is a clear shift in the distribution of ages.
+The increased representation of ages 65+ would likely predict a Trump win.
+Our model coefficient results also indicate that someone being in the oldest age group is more likely to vote for Trump, though significance is only slightly off.
+With an effectively older population compared to the Nationscape polling data, we can conclude that we would be likely see more votes for Trump coming out of our ACS group.
+Similarly, when looking at sex, we observe that there are more women in our ACS dataset than there are men, which is the opposite of the distribution found in the Nationscape data.
+Our model coefficient results again suggest that men are more likely to vote for Trump, and since we have more women in the ACS data, we would expect a higher percentage of Biden votes.
+This is further supported by other research that suggests that Trump has lost favour with white suburban women, a group he previously was well percevied by[@conroy2020].
+A lot of models during the 2016 election season suggested a Clinton win, one of the few models that predicted a Trump win was a model that used MRP[@2016election].
+It is worth mentioning that MRP isn't some sort of magic bullet as a few other MRP models during that time still predicted a Clinton win within the electoral college[@2016election].
+How do current models fair?
+The primary model suggests that Trump has a 91% change of winning, in spite of Trump trailing Biden in the polls[@norpoth_2020].
 
-To begin this discussion, we begin with a quick explanation of the US Federal voting system. This context is important for us to understand some key points and drawbacks with interpreting our model.
-
-Talk about previous research that has been done with similar models and the output of this model.
-Our model seems to indicate that Biden is going to win the popular vote.
-Throw some crap in here about 538 has certain polls that show a slight trump win.
-Show falling trump approvals and other things like that.
-This model has been thoroughly researched and has shown to be effective in multiple other contexts(needs citation)
+Overall, our model predicts a Trump majority.
 
@@
 # References
+ADD FIVE THIRTY EIGHT CITATION
 

### 2020-11-03 scripts/01-data_cleaning-survey.R

@@ -90,7 +90,7 @@ us_ethnicity_mapping <- data.frame("race" = c("white",
 
 
 #
-survey_data_reduced <-survey_data_reduced %>% 
+survey_data_reduced <- survey_data_reduced %>% 
   filter(vote_intention == "Yes, I will vote") %>% # Keep observations of those who intend to vote in the upcoming 2020 election
   filter(vote_2020 == "Donald Trump"| vote_2020 == "Joe Biden") %>% # Since we're only looking for a binary result, we're filtering for respondents that would've voted for Trump or Biden. By nature of the filter conditions, we'd be removing all other types of responses.
   filter(vote_2016 == "Donald Trump"| vote_2016 == "Hillary Clinton")
@@ -125,7 +125,7 @@ saveRDS(survey_data_reduced, file = "inputs/cleaned_data/individual-survey.rds")
 
 ### GRAPHING ###
 
-#these graph were done as EDA to see if there were any notable trends
+#these graph were done as EDA to see if
 s <- survey_data_reduced %>% 
   select(vote_2016, race) %>% 
   group_by(race, vote_2016) %>% 

### 2020-11-03 scripts/01-data_cleaning-survey.R
added more references

@@ -125,7 +125,7 @@ saveRDS(survey_data_reduced, file = "inputs/cleaned_data/individual-survey.rds")
 
 ### GRAPHING ###
 
-#these graph were done as EDA to see if
+#these graph were done as EDA to see if there were any trends
 s <- survey_data_reduced %>% 
   select(vote_2016, race) %>% 
   group_by(race, vote_2016) %>% 

### 2020-11-03 outputs/paper/paper.Rmd

--- 2020-11-03 outputs/paper/paper.Rmd
@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: |
-  | The purpose of this analysis is to forecast the results of the upcoming United States 2020 presidential election.
+abstract: The purpose of this analysis is to forecast the results of the upcoming United States 2020 presidential election.
 In this paper, we first consider the Democracy Fund + UCLA Nationscape Wave 50 dataset, which contains the results of a survey (conducted June 25-July 1, 2020) on American voter attitudes.

### 2020-11-03 outputs/paper/paper.Rmd
more merge issues

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: The purpose of this analysis is to forecast the results of the upcoming United States 2020 presidential election.
+abstract: |
+  | The purpose of this analysis is to forecast the results of the upcoming United States 2020 presidential election.
 In this paper, we first consider the Democracy Fund + UCLA Nationscape Wave 50 dataset, which contains the results of a survey (conducted June 25-July 1, 2020) on American voter attitudes.

### 2020-11-03 outputs/paper/paper.Rmd
F

@@
 The Democracy Fund + UCLA Nationscape is "one of the largest public opinion surveys ever conducted -- interviewing people in nearly every county, congressional district, and mid-sized U.S.
-city in the lead up to the 2020 election." \cite{vavreck_tausanovitch_2020} Consequently, interviewing "roughly 6,250 per week" was part of the survey methodology employed by Nationscape.
+city in the lead up to the 2020 election." @citevavreck_tausanovitch_2020] Consequently, interviewing "roughly 6,250 per week" was part of the survey methodology employed by Nationscape.
 As such, the Nationscape Wave 50 dataset contains results for 6,479 interviews conducted during the week of June 25-July 1, 2020.
-It can be accessed by visiting https://www.voterstudygroup.org/publication/nationscape-data-set and submitting a request to access the data \cite{vavreck_tausanovitch_2020}.
+It can be accessed by visiting https://www.voterstudygroup.org/publication/nationscape-data-set and submitting a request to access the data [@vavreck_tausanovitch_2020].
 By providing your full name and email address, you will then be emailed a link to the page where it can be downloaded.
@@
 Nationscape samples are provided by Lucid, a market research platform that uses their software for conducting online surveys.
-The Nationscape surveys were completed by participants online in English \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019} .
+The Nationscape surveys were completed by participants online in English [tausanovitch_vavreck_reny_hayes_rudkin_2019].
 The survey data is then weighted using a simple raking technique, where one set of weights is generated for each week's survey, in an attempt to have data that is representative of their target population.
@@
 This high quality of survey responses can be at least partially attributed to the fact that "Nationscape samples are provided by Lucid, a market research platform that runs an online exchange for survey respondents.
-\cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}" What is more, the individuals belonging to the sample are required to "complete an attention check before completing the [online] survey \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}." With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
+[@tausanovitch_vavreck_reny_hayes_rudkin_2019]" What is more, the individuals belonging to the sample are required to "complete an attention check before completing the [online] survey [@tausanovitch_vavreck_reny_hayes_rudkin_2019]." With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
 
@@
 While the goal of the Nationscape survey is to determine American voter attitudes, making inferences based only on the Wave 50 dataset would result in sample bias.
-This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match "a set of demographic criteria \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}." It is unlikely that a survey sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
+This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match "a set of demographic criteria [@tausanovitch_vavreck_reny_hayes_rudkin_2019]." It is unlikely that a survey sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
 This may lead one to wonder why Nationscape chose to use the described sampling approach.
@@
 Nonetheless, this is a much smaller percentage of non-response in household income than is typically seen in political surveys, and Nationscape handles it by not weighting income for non-respondents.
-Similarly, a statement from Nationscape in its "Representativeness Assessment" implies an overall lack of concern for non-response in the data, as "previous evaluations of the samples Lucid provides have found them to be of high quality \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}."
+Similarly, a statement from Nationscape in its "Representativeness Assessment" implies an overall lack of concern for non-response in the data, as "previous evaluations of the samples Lucid provides have found them to be of high quality [@tausanovitch_vavreck_reny_hayes_rudkin_2019]."
 
@@
 This allowed for us to look at how age groups might influence the vote.
-The 2020 election marks the first time that Millennials and Gen Z will have an equal share of the number of eligible voters as Baby Boomers and earlier generations \cite{@agearticle} .
+The 2020 election marks the first time that Millennials and Gen Z will have an equal share of the number of eligible voters as Baby Boomers and earlier generations [@agearticle] .
 Older voters in the past have had higher turn out rates, so looking at how how age categories tend to vote can give insight on how this will play a factor in the future election.
@@
 In the 2020 election, Latinos will be the largest racial/ethnic minority in the electoral.
-With Trump's previous comments on countries such as Mexico, it seems important to see how pivotal the vote of this group will be in the election \cite{@citeBBC}.
+With Trump's previous comments on countries such as Mexico, it seems important to see how pivotal the vote of this group will be in the election [@citeBBC].
 
@@
 Certain states have been dominated by one particular party.
-For example, the Democratic party hasn't won the majority vote in the presidential election since 1992 \cite{{@cite270towin}.
+For example, the Democratic party hasn't won the majority vote in the presidential election since 1992 [@cite270towin].
 Since the election is based on electoral colleges as well, the number of Sates won by a candidate should give us a clear prospect for the dominating candidate.
@@
 In order to forecast the popular vote of the U.S 2020 elections, multilevel logistic regression with post-stratification (MRP) was used.
-MRP is useful for when generalizing from a possibly non-representative poll \cite{kennedy2020know}.
-The individual survey gives sample data on voters of the general U.S population, and post-stratification allows us to re-weight estimates, adjusting bias between our sample and the target population, so that we have a representative sample of likely voters \cite{alexander2020a}.
+MRP is useful for when generalizing from a possibly non-representative poll {kennedy2020know}.
+The individual survey gives sample data on voters of the general U.S population, and post-stratification allows us to re-weight estimates, adjusting bias between our sample and the target population, so that we have a representative sample of likely voters {alexander2020a}.
 To achieve this, cells are constructed using variables in the cleaned Nationscape survey such as age, household income, and race.
 The model is then trained on the survey using the proportions given by the  ACS post-stratification data set.
-This approach gives us an advantage when attempted to forecast voting, as we can use a broad survey to speak to subsets in the population, and also tends to be less expensive to collect than non-probability samples \cite{alexander2020a}.
+This approach gives us an advantage when attempted to forecast voting, as we can use a broad survey to speak to subsets in the population, and also tends to be less expensive to collect than non-probability samples {alexander2020a}.
 However, using this approach places limitations on how much we can interpret from the estimates.
@@
 $$ Pr(y_i = 1) = logit^-1(X_i *\beta) $$
-Equation from: \cite{@RAOS}
+Equation from: {@RAOS}
 
@@
 
-All our analysis was done using R \cite{@CiteR}.
+All our analysis was done using R {@CiteR}.
 

### 2020-11-03 outputs/paper/paper.Rmd
commit

--- 2020-11-03 outputs/paper/paper.Rmd
@@
 
-One of the key things we have to explain here is how the electoral college works.
-Readers may remember the 2016 US election where Hillary Clinton won the popular vote, however, this didn't mean that she won the federal election.
-This was due to her not winning the electoral college.
-The electoral college is a body that is elected to determine who will be the next president of the US.
-This happened in 2000 when Bush didn't win the popular vote, but did end up winning the electoral college
-## Second discussion point
-
-## Third discussion point
-
 ## Weaknesses and next steps

### 2020-11-03 outputs/paper/paper.tex
commit

--- 2020-11-03 outputs/paper/paper.Rmd
@@
----
-title: "Prediction of the 2020 Presidential Election"
-subtitle: "Spoiler Alert: Trump Wins"
-author: "Anees Shaikh, Jaffa Romain, Lu Mu, and Cameron Fryer"
-thanks: "Code and data are available at: https://github.com/aneesshake/elections_prediction"
-date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: |
-  | The purpose of this analysis is to forecast the results of the upcoming United States 2020 presidential election.
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Prediction of the 2020 Presidential Election},
+  pdfauthor={Anees Shaikh, Jaffa Romain, Lu Mu, and Cameron Fryer},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\usepackage{longtable,booktabs}
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
+
+\title{Prediction of the 2020 Presidential Election\thanks{Code and data are available at: \url{https://github.com/aneesshake/elections_prediction}}}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{Spoiler Alert: Trump Wins}
+\author{Anees Shaikh, Jaffa Romain, Lu Mu, and Cameron Fryer}
+\date{02 November 2020}
+
+\begin{document}
+\maketitle
+\begin{abstract}
+The purpose of this analysis is to forecast the results of the upcoming United States 2020 presidential election.
 In this paper, we first consider the Democracy Fund + UCLA Nationscape Wave 50 dataset, which contains the results of a survey (conducted June 25-July 1, 2020) on American voter attitudes.
@@
 The model is then applied to the post-stratification dataset; namely, the results of the 2018 1-year American Community Survey (ACS), to get predictions representative of the target population.
-Since the ACS data pertains to individual persons and their characteristics, the associated use of our model allows us to conclude that Donald Trump will be relected as the President of the United States, winning the popular vote by ___ with a margin of error of ______
-Keywords: forecasting; US 2020 election; Trump; Biden; multilevel regression with post-stratification
-output:
-  bookdown::pdf_document2:
-    citation_package: natbib
-toc: FALSE
-bibliography: references.bib
----
+Since the ACS data pertains to individual persons and their characteristics, the associated use of our model allows us to conclude that Donald Trump will be relected as the President of the United States, winning the popular vote by \_\_\_ with a margin of error of \_\_\_\_\_\_
+\end{abstract}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-library(tidyverse)
-library(tinytex)
-library(readr)
-library(here)
-library(dplyr)
-library(ggplot2)
-library(ggthemes)
-library(ResourceSelection)
-library(maps)
-library(mapproj)
-library(caret)
-library(statebins)
-```
+\ldots..
 
-# Introduction
+\hypertarget{data}{%
+\section{Data}\label{data}}
 
-.....
-
-# Data
-
- ## Individual Survey Data
+\#\# Individual Survey Data
 
@@
 To forecast the results of the presidential election, our model was fitted using the Nationscape survey data.
-The Democracy Fund + UCLA Nationscape is "one of the largest public opinion surveys ever conducted -- interviewing people in nearly every county, congressional district, and mid-sized U.S.
-city in the lead up to the 2020 election." \cite{vavreck_tausanovitch_2020} Consequently, interviewing "roughly 6,250 per week" was part of the survey methodology employed by Nationscape.
+The Democracy Fund + UCLA Nationscape is ``one of the largest public opinion surveys ever conducted -- interviewing people in nearly every county, congressional district, and mid-sized U.S.
+city in the lead up to the 2020 election.'' \cite{vavreck_tausanovitch_2020} Consequently, interviewing ``roughly 6,250 per week'' was part of the survey methodology employed by Nationscape.
 As such, the Nationscape Wave 50 dataset contains results for 6,479 interviews conducted during the week of June 25-July 1, 2020.
-It can be accessed by visiting https://www.voterstudygroup.org/publication/nationscape-data-set and submitting a request to access the data \cite{vavreck_tausanovitch_2020}.
+It can be accessed by visiting \url{https://www.voterstudygroup.org/publication/nationscape-data-set} and submitting a request to access the data \cite{vavreck_tausanovitch_2020}.
 By providing your full name and email address, you will then be emailed a link to the page where it can be downloaded.
@@
 Another strength of this dataset is the overall quality of responses from those surveyed.
-More specifically, the data does not contain any duplicate records, inappropriate entries, or poor entries (e.g.
-typos, misspellings); hence, making it easier to work with.
-This high quality of survey responses can be at least partially attributed to the fact that "Nationscape samples are provided by Lucid, a market research platform that runs an online exchange for survey respondents.
-\cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}" What is more, the individuals belonging to the sample are required to "complete an attention check before completing the [online] survey \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}." With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
+More specifically, the data does not contain any duplicate records, inappropriate entries, or poor entries (e.g.~typos, misspellings); hence, making it easier to work with.
+This high quality of survey responses can be at least partially attributed to the fact that ``Nationscape samples are provided by Lucid, a market research platform that runs an online exchange for survey respondents.
+\cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}'' What is more, the individuals belonging to the sample are required to ``complete an attention check before completing the {[}online{]} survey \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}.'' With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
 
@@
 While the goal of the Nationscape survey is to determine American voter attitudes, making inferences based only on the Wave 50 dataset would result in sample bias.
-This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match "a set of demographic criteria \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}." It is unlikely that a survey sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
+This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match ``a set of demographic criteria \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}.'' It is unlikely that a survey sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
 This may lead one to wonder why Nationscape chose to use the described sampling approach.
@@
 For example, although there is potential coverage error in the data--a common result of online surveys--some trade-offs are a lesser rate of non-response, lower cost, convenience and accuracy.
-Whilst on the topic of non-response, it is worth mentioning that the variable with greatest amount of non-response in the Wave 50 data set is household income at roughly 5.5%.
+Whilst on the topic of non-response, it is worth mentioning that the variable with greatest amount of non-response in the Wave 50 data set is household income at roughly 5.5\%.
 Nonetheless, this is a much smaller percentage of non-response in household income than is typically seen in political surveys, and Nationscape handles it by not weighting income for non-respondents.
-Similarly, a statement from Nationscape in its "Representativeness Assessment" implies an overall lack of concern for non-response in the data, as "previous evaluations of the samples Lucid provides have found them to be of high quality \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}."
+Similarly, a statement from Nationscape in its ``Representativeness Assessment'' implies an overall lack of concern for non-response in the data, as ``previous evaluations of the samples Lucid provides have found them to be of high quality \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}.''
 
@@
 
-```{r ageind, echo=FALSE, fig.cap="Distribution of Age in Nationscape data"}
-individual_survey <- readRDS(here::here("inputs/cleaned_data","individual-survey.rds"))
-post_strat <- readRDS(here::here("inputs/cleaned_data","post-strat.rds"))
-
-cont <- mutate(individual_survey, vote = (case_when(individual_survey$vote_2020 == 1~"Trump", individual_survey$vote_2020 == 0~"Biden")))
-
-  ggplot(cont, aes(x = age, fill  = vote), color = "black") + 
-  geom_bar(alpha = 0.7, color = "black") + 
-  scale_fill_manual(values = c("#0000FF","#FF4040")) + 
-  labs(title = "Age Vote Distribution" , 
-       subtitle =  " A look into how eligible voters intend to vote based on age.", 
-       x = "Age", 
-       y= "Number of votes", 
-       fill = "Nominee") +   
-  theme(axis.text.x = element_text(angle = 35, hjust = 1, size = 5))
-
-```
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/ageind-1.pdf}
+\caption{\label{fig:ageind}Distribution of Age in Nationscape data}
+\end{figure}
 
@@
 Older voters in the past have had higher turn out rates, so looking at how how age categories tend to vote can give insight on how this will play a factor in the future election.
-We can see in (Figure \@ref(fig:ageind)) that there is a much smaller proportion of younger voters, but this is the age group that seems to have the strongest Democratic support.
+We can see in (Figure \ref{fig:ageind}) that there is a much smaller proportion of younger voters, but this is the age group that seems to have the strongest Democratic support.
 
@@
 
-```{r racedis, echo=FALSE, fig.cap= "Votes by Race in Nationscape data"}
-#these graph were done as EDA to see if there were any notable trends
-s <-cont %>% 
-  select(vote_2016, race) %>% 
-  group_by(race, vote_2016) %>% 
-  summarise(n = n()) %>% 
-  mutate(freq = n/sum(n))
-## 2016 vote by race
-s %>% ggplot(aes(x = vote_2016, fill = race, group = race)) + 
-  geom_bar(aes(y = freq), stat = "identity", position = "dodge", alpha = 2/3) + 
-  labs(x = "Race", title = " Percentage of 2016 Presidential Election Votes by Race")
-```
+\begin{verbatim}
+## `summarise()` regrouping output by 'race' (override with `.groups` argument)
+\end{verbatim}
 
-Race: (Figure \@ref(fig:racedis)) shows the distribution of votes in the 2016 federal election; specifically the Democratic candidate, Hillary Clinton, and the Republican candidate, Donald Trump.
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/racedis-1.pdf}
+\caption{\label{fig:racedis}Votes by Race in Nationscape data}
+\end{figure}
+
+Race: (Figure \ref{fig:racedis}) shows the distribution of votes in the 2016 federal election; specifically the Democratic candidate, Hillary Clinton, and the Republican candidate, Donald Trump.
 We can see that certain groups had a significant different in their voting preference.
@@
 Certain states have been dominated by one particular party.
-For example, the Democratic party hasn't won the majority vote in the presidential election since 1992 \cite{{@cite270towin}.
+For example, the Democratic party hasn't won the majority vote in the presidential election since 1992 \textbackslash{}\href{mailto:cite\%7B\%7B@cite270towin}{\nolinkurl{cite\{\{@cite270towin}}\}.
 Since the election is based on electoral colleges as well, the number of Sates won by a candidate should give us a clear prospect for the dominating candidate.
 
- ## Post-stratification Data
+\#\# Post-stratification Data
 
 For our next step in our analysis, we use post-stratification to overcome any bias and have data representative of our target population.
-Conducted on a yearly basis, the American Community Survey "is a nationwide survey that collects and produces information on social, economic, housing, and demographic characteristics" of the United States population.
-In order to do so, the United States Census Bureau "mails questionnaires to approximately 295,000 addresses a month across the [country]." It should be noted that all the addresses receiving a letter to participate in the ACS are from a random sample chosen by the Census Bureau each month.
+Conducted on a yearly basis, the American Community Survey ``is a nationwide survey that collects and produces information on social, economic, housing, and demographic characteristics'' of the United States population.
+In order to do so, the United States Census Bureau ``mails questionnaires to approximately 295,000 addresses a month across the {[}country{]}.'' It should be noted that all the addresses receiving a letter to participate in the ACS are from a random sample chosen by the Census Bureau each month.
 Furthermore, every single address in the United States has approximately a 1 in 480 chance of being randomly selected to participate in the ACS during a given month.
 The exception to this comes from addresses that have already been selected to participate in the last 5 years; they are not allowed to be selected for another monthly sample.
-  To access the 2018 1-year ACS dataset, the first step is to visit https://usa.ipums.org/usa/ and create an account.
+To access the 2018 1-year ACS dataset, the first step is to visit \url{https://usa.ipums.org/usa/} and create an account.
 From there, IPUMS will send you an email to verify your email address.
@@
 Once on the website, you will be able to access the page where ACS sample and variables can be selected.
-To choose the 2018 1-year ACS as your sample, press "Select Samples".
+To choose the 2018 1-year ACS as your sample, press ``Select Samples''.
 For this analysis, the single 2018 sample was chosen.
@@
 The variables that we selected for the dataset used in this paper are: region, stateicp, hhincome, sex, age, marst, race, hispan, bpl, citizen, educ, labforce, and ftotinc.
-When finished selecting variables, click "View Cart" near the top right of the page, and you will arrive at the cart page.
+When finished selecting variables, click ``View Cart'' near the top right of the page, and you will arrive at the cart page.
 This is your chance to look over all of the chosen variables.
-If you are happy with them, press the blue button near the middle of the page that reads "Create Data Extract." On the new page that you arrive at, there is a blue button near the bottom of the page with "Customize Sample Sizes" displayed in it.
+If you are happy with them, press the blue button near the middle of the page that reads ``Create Data Extract.'' On the new page that you arrive at, there is a blue button near the bottom of the page with ``Customize Sample Sizes'' displayed in it.
 If you wish to change the size of your sample, then click the button.
-Now that our dataset has been fully customized, press the blue "Submit Extract" button at the very bottom of the page.
+Now that our dataset has been fully customized, press the blue ``Submit Extract'' button at the very bottom of the page.
 Finally, you have reached the page from which your dataset can be downloaded, and the final step is to wait for the sample to be ready.
@@
 The sample itself includes all Americans living at one of the 295,000 addresses that are selected to participate in the ACS during a given month.
-Furthermore, the Census Bureau designs the sample in a way that guarantees favourable geographic inclusion, while also utilizing random sampling to select the "addresses to be included in the ACS." Every resident of the addresses which the invitation to complete the ACS was mailed is required to complete either the online survey, or "to mail the completed paper questionnaire." If neither requirement is satisfied after about a month, the Census Bureau "will mail an additional paper survey questionnaire." If the second paper questionnaire goes uncompleted too, then roughly 1 month after it was sent, someone will show up at your address to conduct an in-person interview with you.
+Furthermore, the Census Bureau designs the sample in a way that guarantees favourable geographic inclusion, while also utilizing random sampling to select the ``addresses to be included in the ACS.'' Every resident of the addresses which the invitation to complete the ACS was mailed is required to complete either the online survey, or ``to mail the completed paper questionnaire.'' If neither requirement is satisfied after about a month, the Census Bureau ``will mail an additional paper survey questionnaire.'' If the second paper questionnaire goes uncompleted too, then roughly 1 month after it was sent, someone will show up at your address to conduct an in-person interview with you.
 Due to the fact that completing the ACS is the law in the United States, non-response does not appear to be a very big issue.
@@
 
-# Model
+\hypertarget{model}{%
+\section{Model}\label{model}}
 
@@
 As we cannot take into consideration any other candidates running in the election due to the binary outcome.
-The variable 'vote_2020' in the data set was used as the dependent variable (the variable we are trying to predict), with the variable returning 1 for voters intending to vote for Trump, and 0 if they are voting for Biden.
+The variable `vote\_2020' in the data set was used as the dependent variable (the variable we are trying to predict), with the variable returning 1 for voters intending to vote for Trump, and 0 if they are voting for Biden.
 
@@
 The model takes the form:
-$$ Pr(y_i = 1) = logit^-1(X_i *\beta) $$
+\[ Pr(y_i = 1) = logit^-1(X_i *\beta) \]
 Equation from: \cite{@RAOS}
 
-where $y_i = 1$ represents a voter voting for Trump in the election.
-$Pr(y_i = 1)$ is the probability of a voter choosing to vote for Trump.
-The $X_i\beta$ are the linear predictors, where $\beta$ represents the fitted coefficients for each independent variable $X_i$ in the model.
-The coefficient values signify the mean variable changes in a one unit shift in the given variable $\beta$.
+where \(y_i = 1\) represents a voter voting for Trump in the election.
+\(Pr(y_i = 1)\) is the probability of a voter choosing to vote for Trump.
+The \(X_i\beta\) are the linear predictors, where \(\beta\) represents the fitted coefficients for each independent variable \(X_i\) in the model.
+The coefficient values signify the mean variable changes in a one unit shift in the given variable \(\beta\).
 
-The logit function $logit(x) = log(x/1-x)$ maps the range (0, 1) to $(-\infty, \infty)$.
-Its inverse function $logit^-1(x) = e^x/(1+e^x)$ maps back to the unit range. 
+The logit function \(logit(x) = log(x/1-x)\) maps the range (0, 1) to \((-\infty, \infty)\).
+Its inverse function \(logit^-1(x) = e^x/(1+e^x)\) maps back to the unit range.
 The model's output is bounded between 0 and 1, mapping the outputs into a binary outcome.
-The output tells us the probability that some person $i$, will vote for Trump depends on their age, sex, state, household income, and whether the voter is Hispanic.
+The output tells us the probability that some person \(i\), will vote for Trump depends on their age, sex, state, household income, and whether the voter is Hispanic.
 
-We fit our regression model using the `glm()` function in [`R`].  We can see from table 1, which was made using `kable` from `knitr`,  that the fitted model results in the following coefficients:
-```{r, echo=FALSE}
-# loading data sets and cell counts for MRP
-cell_counts <- readRDS(here::here("inputs/cleaned_data", "cell_counts.rds"))
+We fit our regression model using the \texttt{glm()} function in {[}\texttt{R}{]}.
+We can see from table 1, which was made using \texttt{kable} from \texttt{knitr}, that the fitted model results in the following coefficients:
 
-#Loading in the final model and the coefficents of the model in this cell.
-Please note that the outputs of those models are being stored in the outputs/model directory.
-The here package should give you the same result but if it doesn't then you may have to manually enter the same directory.
+\begin{table}
 
-final_model <- readRDS(here::here("outputs/model","final_model.rds"))
-coefs <- readRDS(here::here("outputs/model","coefficients.rds"))
-knitr::kable(coefs, caption = " Table 1: Model Coefficients", digits = 2)
+\caption{\label{tab:unnamed-chunk-1} Table 1: Model Coefficients}
+\centering
+\begin{tabular}[t]{l|r|r|r|r}
+\hline
+term & estimate & std.error & statistic & p.value\\
+\hline
+(Intercept) & 1.52 & 0.71 & 2.15 & 0.03\\
+\hline
+sexmale & 0.40 & 0.07 & 5.45 & 0.00\\
+\hline
+age25-44 & 0.35 & 0.20 & 1.73 & 0.08\\
+\hline
+age45-64 & 0.46 & 0.21 & 2.22 & 0.03\\
+\hline
+age65+ & 0.38 & 0.21 & 1.81 & 0.07\\
+\hline
+raceblack & -3.13 & 0.45 & -6.98 & 0.00\\
+\hline
+racechinese & -2.29 & 0.59 & -3.87 & 0.00\\
+\hline
+racejapanese & -1.97 & 0.79 & -2.50 & 0.01\\
+\hline
+raceother asian or pacific islander & -1.38 & 0.47 & -2.95 & 0.00\\
+\hline
+raceother race, nec & -1.07 & 0.44 & -2.42 & 0.02\\
+\hline
+racewhite & -0.55 & 0.41 & -1.33 & 0.18\\
+\hline
+hispanmexican & -1.86 & 0.57 & -3.26 & 0.00\\
+\hline
+hispannot hispanic & -1.22 & 0.55 & -2.22 & 0.03\\
+\hline
+hispanother & -1.45 & 0.58 & -2.51 & 0.01\\
+\hline
+hispanpuerto rican & -1.73 & 1.47 & -1.18 & 0.24\\
+\hline
+\end{tabular}
+\end{table}
 
-```
 In logistic regression, collinearity implies that we have predictor variables that are highly correlated; that is, they have a linear relationship.
@@
 We can see in table 2 below that all VIF values are less than 5, so collinearity will not be a concern for our model.
-```{r, echo=FALSE}
-cor <- readRDS(here::here("outputs/model","cor.rds"))
-knitr::kable(cor, caption =  "Checking Model Collinearity", digits = 2)
-```
 
-```{r, echo=FALSE}
- htst <- hoslem.test(final_model$y, fitted(final_model), g=10)
-htst
-```
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-2}Checking Model Collinearity}
+\centering
+\begin{tabular}[t]{l|r|r|r}
+\hline
+  & GVIF & Df & GVIF\textasciicircum{}(1/(2*Df))\\
+\hline
+sex & 1.08 & 1 & 1.04\\
+\hline
+age & 1.24 & 3 & 1.04\\
+\hline
+race & 1.64 & 6 & 1.04\\
+\hline
+household\_income & 1.81 & 23 & 1.01\\
+\hline
+hispan & 1.43 & 4 & 1.05\\
+\hline
+state & 2.24 & 50 & 1.01\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{verbatim}
+## 
+##  Hosmer and Lemeshow goodness of fit (GOF) test
+## 
+## data:  final_model$y, fitted(final_model)
+## X-squared = 3.5112, df = 8, p-value = 0.8983
+\end{verbatim}
+
 Before the model can be used for any predictions, we must also check that our model fits our data well, and meets any assumptions made by the model.
@@
 
-| Table 4: Hosmer and Lemeshow goodness of fit (GOF) test 	|    	|         	|
-|---------------------------------------------------------	|----	|---------	|
-| X-squared                                               	| df 	| p-value 	|
-| 3.5                                                     	| 8  	| 0.9     	|
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+Table 4: Hosmer and Lemeshow goodness of fit (GOF) test & &\tabularnewline
+\midrule
+\endhead
+X-squared & df & p-value\tabularnewline
+3.5 & 8 & 0.9\tabularnewline
+\bottomrule
+\end{longtable}
 
-```{r, echo=FALSE}
-# make predictions using our model with data from post-strat data set
+\begin{verbatim}
+## # A tibble: 7,131 x 8
+##    state_name  sex   age   race          hispan  num_records    prop state_count
+##    <fct>       <fct> <chr> <chr>         <fct>         <int>   <dbl>       <int>
+##  1 connecticut male  18-24 american ind~ not hi~           1 4.00e-5       24986
+##  2 connecticut male  18-24 american ind~ puerto~           2 8.00e-5       24986
+##  3 connecticut male  18-24 american ind~ other             1 4.00e-5       24986
+##  4 connecticut male  18-24 black         not hi~          89 3.56e-3       24986
+##  5 connecticut male  18-24 black         puerto~           4 1.60e-4       24986
+##  6 connecticut male  18-24 chinese       not hi~           5 2.00e-4       24986
+##  7 connecticut male  18-24 japanese      not hi~           1 4.00e-5       24986
+##  8 connecticut male  18-24 other asian ~ not hi~          31 1.24e-3       24986
+##  9 connecticut male  18-24 other race, ~ mexican           8 3.20e-4       24986
+## 10 connecticut male  18-24 other race, ~ puerto~          20 8.00e-4       24986
+## # ... with 7,121 more rows
+\end{verbatim}
 
-cell_counts
+\begin{verbatim}
+## `summarise()` ungrouping output (override with `.groups` argument)
+\end{verbatim}
 
-cell_counts$estimate <- predict(final_model, newdata = cell_counts, type = "response")
-#  prediction intervals
-# fitted value - error to get margin of error
-se_md <- predict(final_model, newdata = cell_counts, type = "response", se.fit = T)
-
- lwr <-  cell_counts$estimate - 1.96*se_md$se.fit
- upper <- cell_counts$estimate + 1.96*se_md$se.fit
-
-# all estimates with lower and upper bounds
-estimates_with_margins <- cell_counts %>% mutate(n = num_records * estimate) %>% cbind(cell_counts, lwr, upper)
-# removing duplicate column name estimate
-estimates_with_margins <- estimates_with_margins[, !duplicated(colnames(estimates_with_margins))]
-
-# post-stratified estimates - number of trump votes with prediction intervals
-post_strat_estimates <- estimates_with_margins %>% 
-  mutate(predict_prop = prop * estimate, l_predict = lwr * prop, u_predict = upper * prop) %>% 
-  group_by(state_name) %>% 
-  summarise(predicted_vote = sum(predict_prop), lower = sum(l_predict), upp = sum(u_predict)) %>% group_by(state_name) %>% ungroup()
-
-post_strat_estimates <- post_strat_estimates %>% 
-  mutate(biden_vote_proportion = 1-predicted_vote,
-         conservative_election_status = if_else(lower %>% between(0.48,0.52),"toss-up",
-                                                 if_else(lower > 0.52, "win","loss")),
-         liberal_election_status  = if_else(upp %>% between(0.48,0.52),"toss-up",
-                                                 if_else(upp > 0.52, "win","loss")),
-         middle_ground_election_status = if_else(predicted_vote %>% between(0.48,0.52),"toss-up",
-                                                 if_else(predicted_vote > 0.52, "win","loss")))
-
-```
-
-# Results
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
@@
 
-## comparison on age
-```{r age, echo=FALSE, include=FALSE, fig.cap="Distribution of Age between ACS and Nationscape data"}
-individual_age_dist<- individual_survey %>% 
-  group_by(age) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(percentage = 100 * num_records/sum(num_records),
-         total_num = sum(num_records),
-         data_source = "Nationscape")
+\hypertarget{comparison-on-age}{%
+\subsection{comparison on age}\label{comparison-on-age}}
 
-post_strat_age_dist <- post_strat %>% 
-  group_by(age) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(percentage = 100 * num_records/sum(num_records),
-         total_num = sum(num_records),
-         data_source = "ACS")
-
-individual_age_dist %>% 
-  union(post_strat_age_dist) %>% 
-  ggplot(aes(x = age, y = percentage)) + 
-  geom_col() + 
-  facet_grid(~data_source) + 
-  theme_economist_white()
-```
-Looking at (Figure \@ref(fig:age)), we observe a shift in age distributions between the two datasets.
+Looking at (Figure \ref{fig:age}), we observe a shift in age distributions between the two datasets.
 There is a slight peak in the 25-44 age group in the nationscape data.
@@
 
-## comparison on sex
+\hypertarget{comparison-on-sex}{%
+\subsection{comparison on sex}\label{comparison-on-sex}}
 
-```{r}
-individual_sex_dist<- individual_survey %>% 
-  group_by(sex) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(percentage = 100 * num_records/sum(num_records),
-         total_num = sum(num_records),
-         data_source = "Nationscape")
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{individual_sex_dist<-}\StringTok{ }\NormalTok{individual_survey }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{group_by}\NormalTok{(sex) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{num_records =} \KeywordTok{n}\NormalTok{()) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{percentage =} \DecValTok{100} \OperatorTok{*}\StringTok{ }\NormalTok{num_records}\OperatorTok{/}\KeywordTok{sum}\NormalTok{(num_records),}
+         \DataTypeTok{total_num =} \KeywordTok{sum}\NormalTok{(num_records),}
+         \DataTypeTok{data_source =} \StringTok{"Nationscape"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-post_strat_sex_dist <- post_strat %>% 
-  group_by(sex) %>% 
-  summarise(num_records = n()) %>% 
-  mutate(percentage = 100 * num_records/sum(num_records),
-         total_num = sum(num_records),
-         data_source = "ACS")
+\begin{verbatim}
+## `summarise()` ungrouping output (override with `.groups` argument)
+\end{verbatim}
 
-individual_sex_dist %>% 
-  union(post_strat_sex_dist) %>% 
-  ggplot(aes(x = sex, y = percentage)) + 
-  geom_col() + 
-  facet_grid(~data_source) + 
-  theme_economist_white()
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{post_strat_sex_dist <-}\StringTok{ }\NormalTok{post_strat }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{group_by}\NormalTok{(sex) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{num_records =} \KeywordTok{n}\NormalTok{()) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{percentage =} \DecValTok{100} \OperatorTok{*}\StringTok{ }\NormalTok{num_records}\OperatorTok{/}\KeywordTok{sum}\NormalTok{(num_records),}
+         \DataTypeTok{total_num =} \KeywordTok{sum}\NormalTok{(num_records),}
+         \DataTypeTok{data_source =} \StringTok{"ACS"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-```
+\begin{verbatim}
+## `summarise()` ungrouping output (override with `.groups` argument)
+\end{verbatim}
 
-```{r sex_vote,fig.cap="Which individual did the candidate vote for? 2020 on the left, 2016 on the right"}
-#figure to see 2020 voting by sex
-vote_2020_sex <- individual_survey %>% 
-  mutate(vote_2020 = if_else(vote_2020 == 1,"Donald Trump","Joe Biden")) %>% 
-  group_by(sex,vote_2020) %>% 
-  tally() %>% 
-  ggplot(aes(x=sex, y = n)) + 
-  geom_col() + 
-  facet_grid(~vote_2020)+ labs(x = "Sex", y = "Number of people")
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{individual_sex_dist }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{union}\NormalTok{(post_strat_sex_dist) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ sex, }\DataTypeTok{y =}\NormalTok{ percentage)) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{geom_col}\NormalTok{() }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{facet_grid}\NormalTok{(}\OperatorTok{~}\NormalTok{data_source) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{theme_economist_white}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
 
-#figure to see 2016 voting by sex
-vote_2016_sex <- individual_survey %>% 
-  group_by(sex,vote_2016) %>% 
-  tally() %>% 
-  ggplot(aes(x=sex, y = n)) + 
-  geom_col() + 
-  facet_grid(~vote_2016) + labs(x = "Sex", y = "Number of people")
+\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}
 
-library(patchwork)
-vote_2020_sex + vote_2016_sex
-```
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{#figure to see 2020 voting by sex}
+\NormalTok{vote_}\DecValTok{2020}\NormalTok{_sex <-}\StringTok{ }\NormalTok{individual_survey }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{vote_2020 =} \KeywordTok{if_else}\NormalTok{(vote_}\DecValTok{2020} \OperatorTok{==}\StringTok{ }\DecValTok{1}\NormalTok{,}\StringTok{"Donald Trump"}\NormalTok{,}\StringTok{"Joe Biden"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{group_by}\NormalTok{(sex,vote_}\DecValTok{2020}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{tally}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{sex, }\DataTypeTok{y =}\NormalTok{ n)) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{geom_col}\NormalTok{() }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{facet_grid}\NormalTok{(}\OperatorTok{~}\NormalTok{vote_}\DecValTok{2020}\NormalTok{)}\OperatorTok{+}\StringTok{ }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x =} \StringTok{"Sex"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Number of people"}\NormalTok{)}
 
-Figure (@fig) indicates that a man is much more likely to vote for Trump than a woman is. This trend is also observed in the 2016 elections when Hillary Clinton, a woman, ran for office.
+\CommentTok{#figure to see 2016 voting by sex}
+\NormalTok{vote_}\DecValTok{2016}\NormalTok{_sex <-}\StringTok{ }\NormalTok{individual_survey }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{group_by}\NormalTok{(sex,vote_}\DecValTok{2016}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{tally}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{sex, }\DataTypeTok{y =}\NormalTok{ n)) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{geom_col}\NormalTok{() }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{facet_grid}\NormalTok{(}\OperatorTok{~}\NormalTok{vote_}\DecValTok{2016}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x =} \StringTok{"Sex"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Number of people"}\NormalTok{)}
 
-```{r survey_election_results}
-individual_survey %>% mutate(vote_2020 = if_else(vote_2020 == 1,"Donald Trump","Joe Biden")) %>% group_by(vote_2020) %>% tally() %>% rename(candidate = vote_2020, votes = n ) %>% knitr::kable(caption = "Polling results for 2020 Election")
-```
+\KeywordTok{library}\NormalTok{(patchwork)}
+\NormalTok{vote_}\DecValTok{2020}\NormalTok{_sex }\OperatorTok{+}\StringTok{ }\NormalTok{vote_}\DecValTok{2016}\NormalTok{_sex}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/sex_vote-1.pdf}
+\caption{(\#fig:sex\_vote)Which individual did the candidate vote for? 2020 on the left, 2016 on the right}
+\end{figure}
+
+Figure (\citet{fig}) indicates that a man is much more likely to vote for Trump than a woman is. This trend is also observed in the 2016 elections when Hillary Clinton, a woman, ran for office.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{individual_survey }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{vote_2020 =} \KeywordTok{if_else}\NormalTok{(vote_}\DecValTok{2020} \OperatorTok{==}\StringTok{ }\DecValTok{1}\NormalTok{,}\StringTok{"Donald Trump"}\NormalTok{,}\StringTok{"Joe Biden"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{group_by}\NormalTok{(vote_}\DecValTok{2020}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{tally}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{candidate =}\NormalTok{ vote_}\DecValTok{2020}\NormalTok{, }\DataTypeTok{votes =}\NormalTok{ n ) }\OperatorTok{%>%}\StringTok{ }\NormalTok{knitr}\OperatorTok{::}\KeywordTok{kable}\NormalTok{(}\DataTypeTok{caption =} \StringTok{"Polling results for 2020 Election"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\textbackslash begin\{table\}
+
+\textbackslash caption\{(\#tab:survey\_election\_results)Polling results for 2020 Election\}
+\centering
+
+\begin{tabular}[t]{l|r}
+\hline
+candidate & votes\\
+\hline
+Donald Trump & 1802\\
+\hline
+Joe Biden & 1739\\
+\hline
+\end{tabular}
+
+\textbackslash end\{table\}
 The above table indicates that Trump is most likely to win. This is captured from Nationscape data.
 
-```{r model2020prediction}
-post_strat_estimates %>% group_by(middle_ground_election_status) %>% tally() %>% rename(state_result = middle_ground_election_status, number = n) %>%  knitr::kable(caption = "Model Prediction for 2020")
-```
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{post_strat_estimates }\OperatorTok{%>%}\StringTok{ }\KeywordTok{group_by}\NormalTok{(middle_ground_election_status) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{tally}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{state_result =}\NormalTok{ middle_ground_election_status, }\DataTypeTok{number =}\NormalTok{ n) }\OperatorTok{%>%}\StringTok{  }\NormalTok{knitr}\OperatorTok{::}\KeywordTok{kable}\NormalTok{(}\DataTypeTok{caption =} \StringTok{"Model Prediction for 2020"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-```{r, echo = FALSE}
+\begin{table}
 
-# estimated proportion of votes by State
-post_strat_estimates %>%
-ggplot(aes(y = predicted_vote, x = state_name, color = "MRP estimate")) + geom_point() +
-geom_errorbar(aes(ymin = lower, ymax = upp), width = 0) +
-ylab("Proportion of Trump Supporters") +
-xlab("State") +
-geom_point(data = individual_survey %>%
-group_by(state_name, vote_2020) %>% summarise(n = n()) %>% group_by(state_name) %>%
-mutate(prop = n/sum(n)) %>% filter(vote_2020==1),
-aes(state_name, prop, color = "Raw data")) + theme_minimal() +
-scale_color_brewer(palette = "Set1") + theme(legend.position = "bottom") + theme(legend.title = element_blank()) + 
-  theme(axis.text.x = element_text(angle = 35, hjust = 1, size = 5)) + coord_flip() +labs(title = "2020 Predicted Popular Vote by State", subtitle = "Comparing predicted popular vote by state using raw data and MRP")
-```
+\caption{\label{tab:model2020prediction}Model Prediction for 2020}
+\centering
+\begin{tabular}[t]{l|r}
+\hline
+state\_result & number\\
+\hline
+loss & 9\\
+\hline
+toss-up & 10\\
+\hline
+win & 32\\
+\hline
+\end{tabular}
+\end{table}
 
-```{r heatmap, echo = FALSE, fig.cap="Heatmap showcasing model predictions at a state-level"}
-# Sorting State wins
-election_wins <- 
-  post_strat_estimates %>% 
-  mutate(winner = if_else(predicted_vote>=0.50,"Trump","Biden")) %>% 
-  group_by(state_name)
+\begin{verbatim}
+## `summarise()` regrouping output by 'state_name' (override with `.groups` argument)
+\end{verbatim}
 
-election_wins$state <- str_to_title(as.character(election_wins$state_name))
+\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
 
-election_wins <- election_wins %>% filter(state != "district of columbia")
+\begin{verbatim}
+## Warning in validate_states(state_data, state_col, merge.x): Found invalid state
+## values: District Of Columbia
+\end{verbatim}
 
-# Plotting state wins
-statebins(election_wins, value_col="winner", ggplot2_scale_function = scale_fill_manual, values = c(Trump ="#b2182b" , Biden = "#2166ac"), name = "state") +
-  theme_statebins(legend_position="right")
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/heatmap-1.pdf}
+\caption{\label{fig:heatmap}Heatmap showcasing model predictions at a state-level}
+\end{figure}
 
-```
+Figure(\ref{fig:heatmap}) shows the predictions of our model on the post-stratified dataset. This figure ignores close calls and has a binary Trump or Biden result.
 
-Figure(\@ref(fig:heatmap)) shows the predictions of our model on the post-stratified dataset. This figure ignores close calls and has a binary Trump or Biden result.
-
-# Discussion
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 with that background knowledge in mind, what do our poll results indicate?
-From Figure(\@ref(fig:survey_election_results)), we gather that Trump has a very slight lead.
+From Figure(@ref(fig:survey\_election\_results)), we gather that Trump has a very slight lead.
 Multiple other polls have shown Biden to be in the lead(FiveThirtyEight, 2020).
 Now there are a few reasons as to why we may not want to use poll results to predict elections.
-Polls have of late garnered a lot of criticism due to their incorrect predictions of the 2016 Federal election results[@citeAapor].
-Note that while they got the ultimate results wrong, they did get the popular vote prediction correct[@citeAapor].Nearly every national poll that year had Hillary Clinton winning, with some estimates of her having a 3% point lead[@citeAapor].
-Ultimately, she won the popular vote by 2.1%.
+Polls have of late garnered a lot of criticism due to their incorrect predictions of the 2016 Federal election results\citep{citeAapor}.
+Note that while they got the ultimate results wrong, they did get the popular vote prediction correct\citep{citeAapor}.Nearly every national poll that year had Hillary Clinton winning, with some estimates of her having a 3\% point lead\citep{citeAapor}.
+Ultimately, she won the popular vote by 2.1\%.
 However, as mentioned previously, the US does not use a popular vote system.
-At the state level, the polls became less effective, indicating a Clinton lead again[@citeAapor].
+At the state level, the polls became less effective, indicating a Clinton lead again\citep{citeAapor}.
 State level polls do translate over to the electoral college a bit more effectively than national polls do.
 Another reason we may not want to overly rely on polls is yet another cautious tale from the 2016 election season.
-In that season, polls had over-represented college-graduates, a group that was more likely to vote for Clinton, thus tipping the scales even more in favour of Clinton[@citeAapor].
+In that season, polls had over-represented college-graduates, a group that was more likely to vote for Clinton, thus tipping the scales even more in favour of Clinton\citep{citeAapor}.
 Capturing this over-representation is key to effectively predicting an election.
@@
 This was done to elucidate the effect of a difference in representation, akin to the difference of representation of educated voters between polled data and actual populations.
-From figure(\@ref(fig:age)), there is a clear shift in the distribution of ages.
+From figure(\ref{fig:age}), there is a clear shift in the distribution of ages.
 The increased representation of ages 65+ would likely predict a Trump win.
@@
 Our model coefficient results again suggest that men are more likely to vote for Trump, and since we have more women in the ACS data, we would expect a higher percentage of Biden votes.
-This is further supported by other research that suggests that Trump has lost favour with white suburban women, a group he previously was well percevied by[@conroy2020].
-A lot of models during the 2016 election season suggested a Clinton win, one of the few models that predicted a Trump win was a model that used MRP[@2016election].
-It is worth mentioning that MRP isn't some sort of magic bullet as a few other MRP models during that time still predicted a Clinton win within the electoral college[@2016election].
+This is further supported by other research that suggests that Trump has lost favour with white suburban women, a group he previously was well percevied by\citep{conroy2020}.
+A lot of models during the 2016 election season suggested a Clinton win, one of the few models that predicted a Trump win was a model that used MRP\citep{2016election}.
+It is worth mentioning that MRP isn't some sort of magic bullet as a few other MRP models during that time still predicted a Clinton win within the electoral college\citep{2016election}.
 How do current models fair?
-The primary model suggests that Trump has a 91% change of winning, in spite of Trump trailing Biden in the polls[@norpoth_2020].
+The primary model suggests that Trump has a 91\% change of winning, in spite of Trump trailing Biden in the polls\citep{norpoth_2020}.
 
@@
 
-## Weaknesses and next steps
+\hypertarget{weaknesses-and-next-steps}{%
+\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}}
 
@@
 
-# Appendix {-}
+\hypertarget{appendix}{%
+\section*{Appendix}\label{appendix}}
+\addcontentsline{toc}{section}{Appendix}
 
@@
 
-# References
+\hypertarget{references}{%
+\section{References}\label{references}}
+
 ADD FIVE THIRTY EIGHT CITATION
 
-Tausanovitch, Chris and Lynn Vavreck. 2020. Democracy Fund + UCLA Nationscape, October 10-17, 2019 (version 20200814). Retrieved from [URL].
+Tausanovitch, Chris and Lynn Vavreck. 2020. Democracy Fund + UCLA Nationscape, October 10-17, 2019 (version 20200814). Retrieved from {[}URL{]}.
 
 Steven Ruggles, Sarah Flood, Ronald Goeken, Josiah Grover, Erin Meyer, Jose Pacas and Matthew Sobek.
-IPUMS USA: Version 10.0 [dataset].
+IPUMS USA: Version 10.0 {[}dataset{]}.
 Minneapolis, MN: IPUMS, 2020.
-https://doi.org/10.18128/D010.V10.0
+\url{https://doi.org/10.18128/D010.V10.0}
+
+  \bibliography{references.bib}
+
+\end{document}

### 2020-11-03 outputs/paper/paper.tex
committing paper

@@
 The Democracy Fund + UCLA Nationscape is ``one of the largest public opinion surveys ever conducted -- interviewing people in nearly every county, congressional district, and mid-sized U.S.
-city in the lead up to the 2020 election.'' \cite{vavreck_tausanovitch_2020} Consequently, interviewing ``roughly 6,250 per week'' was part of the survey methodology employed by Nationscape.
+city in the lead up to the 2020 election.'' \citet{citevavreck_tausanovitch_2020}{]} Consequently, interviewing ``roughly 6,250 per week'' was part of the survey methodology employed by Nationscape.
 As such, the Nationscape Wave 50 dataset contains results for 6,479 interviews conducted during the week of June 25-July 1, 2020.
-It can be accessed by visiting \url{https://www.voterstudygroup.org/publication/nationscape-data-set} and submitting a request to access the data \cite{vavreck_tausanovitch_2020}.
+It can be accessed by visiting \url{https://www.voterstudygroup.org/publication/nationscape-data-set} and submitting a request to access the data \citep{vavreck_tausanovitch_2020}.
 By providing your full name and email address, you will then be emailed a link to the page where it can be downloaded.
@@
 Nationscape samples are provided by Lucid, a market research platform that uses their software for conducting online surveys.
-The Nationscape surveys were completed by participants online in English \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019} .
+The Nationscape surveys were completed by participants online in English {[}tausanovitch\_vavreck\_reny\_hayes\_rudkin\_2019{]}.
 The survey data is then weighted using a simple raking technique, where one set of weights is generated for each week's survey, in an attempt to have data that is representative of their target population.
@@
 This high quality of survey responses can be at least partially attributed to the fact that ``Nationscape samples are provided by Lucid, a market research platform that runs an online exchange for survey respondents.
-\cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}'' What is more, the individuals belonging to the sample are required to ``complete an attention check before completing the {[}online{]} survey \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}.'' With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
+\citep{tausanovitch_vavreck_reny_hayes_rudkin_2019}'' What is more, the individuals belonging to the sample are required to ``complete an attention check before completing the {[}online{]} survey \citep{tausanovitch_vavreck_reny_hayes_rudkin_2019}.'' With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
 
@@
 While the goal of the Nationscape survey is to determine American voter attitudes, making inferences based only on the Wave 50 dataset would result in sample bias.
-This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match ``a set of demographic criteria \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}.'' It is unlikely that a survey sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
+This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match ``a set of demographic criteria \citep{tausanovitch_vavreck_reny_hayes_rudkin_2019}.'' It is unlikely that a survey sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
 This may lead one to wonder why Nationscape chose to use the described sampling approach.
@@
 Nonetheless, this is a much smaller percentage of non-response in household income than is typically seen in political surveys, and Nationscape handles it by not weighting income for non-respondents.
-Similarly, a statement from Nationscape in its ``Representativeness Assessment'' implies an overall lack of concern for non-response in the data, as ``previous evaluations of the samples Lucid provides have found them to be of high quality \cite{tausanovitch_vavreck_reny_hayes_rudkin_2019}.''
+Similarly, a statement from Nationscape in its ``Representativeness Assessment'' implies an overall lack of concern for non-response in the data, as ``previous evaluations of the samples Lucid provides have found them to be of high quality \citep{tausanovitch_vavreck_reny_hayes_rudkin_2019}.''
 
@@
 This allowed for us to look at how age groups might influence the vote.
-The 2020 election marks the first time that Millennials and Gen Z will have an equal share of the number of eligible voters as Baby Boomers and earlier generations \cite{@agearticle} .
+The 2020 election marks the first time that Millennials and Gen Z will have an equal share of the number of eligible voters as Baby Boomers and earlier generations \citep{agearticle} .
 Older voters in the past have had higher turn out rates, so looking at how how age categories tend to vote can give insight on how this will play a factor in the future election.
@@
 In the 2020 election, Latinos will be the largest racial/ethnic minority in the electoral.
-With Trump's previous comments on countries such as Mexico, it seems important to see how pivotal the vote of this group will be in the election \cite{@citeBBC}.
+With Trump's previous comments on countries such as Mexico, it seems important to see how pivotal the vote of this group will be in the election \citep{citeBBC}.
 
@@
 Certain states have been dominated by one particular party.
-For example, the Democratic party hasn't won the majority vote in the presidential election since 1992 \textbackslash{}\href{mailto:cite\%7B\%7B@cite270towin}{\nolinkurl{cite\{\{@cite270towin}}\}.
+For example, the Democratic party hasn't won the majority vote in the presidential election since 1992 \citep{cite270towin}.
 Since the election is based on electoral colleges as well, the number of Sates won by a candidate should give us a clear prospect for the dominating candidate.
@@
 In order to forecast the popular vote of the U.S 2020 elections, multilevel logistic regression with post-stratification (MRP) was used.
-MRP is useful for when generalizing from a possibly non-representative poll \cite{kennedy2020know}.
-The individual survey gives sample data on voters of the general U.S population, and post-stratification allows us to re-weight estimates, adjusting bias between our sample and the target population, so that we have a representative sample of likely voters \cite{alexander2020a}.
+MRP is useful for when generalizing from a possibly non-representative poll \{kennedy2020know\}.
+The individual survey gives sample data on voters of the general U.S population, and post-stratification allows us to re-weight estimates, adjusting bias between our sample and the target population, so that we have a representative sample of likely voters \{alexander2020a\}.
 To achieve this, cells are constructed using variables in the cleaned Nationscape survey such as age, household income, and race.
 The model is then trained on the survey using the proportions given by the ACS post-stratification data set.
-This approach gives us an advantage when attempted to forecast voting, as we can use a broad survey to speak to subsets in the population, and also tends to be less expensive to collect than non-probability samples \cite{alexander2020a}.
+This approach gives us an advantage when attempted to forecast voting, as we can use a broad survey to speak to subsets in the population, and also tends to be less expensive to collect than non-probability samples \{alexander2020a\}.
 However, using this approach places limitations on how much we can interpret from the estimates.
@@
 \[ Pr(y_i = 1) = logit^-1(X_i *\beta) \]
-Equation from: \cite{@RAOS}
+Equation from: \{\citet{RAOS}\}
 
@@
 
-\begin{verbatim}
-## 
-##  Hosmer and Lemeshow goodness of fit (GOF) test
-## 
-## data:  final_model$y, fitted(final_model)
-## X-squared = 3.5112, df = 8, p-value = 0.8983
-\end{verbatim}
-
 Before the model can be used for any predictions, we must also check that our model fits our data well, and meets any assumptions made by the model.
@@
 
-All our analysis was done using R \cite{@CiteR}.
+All our analysis was done using R \{\citet{CiteR}\}.
 
@@
 
-Overall, our model predicts a Trump majority.
+Overall, our model predicts a Trump majority. The
 

### 2020-11-03 outputs/paper/paper.Rmd

@@
  htst <- hoslem.test(final_model$y, fitted(final_model), g=10)
-htst
 ```

### 2020-11-03 outputs/paper/paper.Rmd
..

@@
 
-.....
-
+There are a vast number of ways in which statistical analysis can be sued to foretell future political future  political appointment.
+With that being said, however, the statistical methods one chooses to exploit are usually dependent on the data in question.
+As such, it is important to both understand, and be able to describe the data before considering the optimal approach in working with it.
+By staying true to this procedure deriving inferences about some given
 # Data
@@
 The Democracy Fund + UCLA Nationscape is "one of the largest public opinion surveys ever conducted -- interviewing people in nearly every county, congressional district, and mid-sized U.S.
-city in the lead up to the 2020 election." @citevavreck_tausanovitch_2020] Consequently, interviewing "roughly 6,250 per week" was part of the survey methodology employed by Nationscape.
+city in the lead up to the 2020 election (Tausanovitch, 2020)." Consequently, interviewing "roughly 6,250 per week" was part of the survey methodology employed by Nationscape.
 As such, the Nationscape Wave 50 dataset contains results for 6,479 interviews conducted during the week of June 25-July 1, 2020.
-It can be accessed by visiting https://www.voterstudygroup.org/publication/nationscape-data-set and submitting a request to access the data [@vavreck_tausanovitch_2020].
+It can be accessed by visiting https://www.voterstudygroup.org/publication/nationscape-data-set and submitting a request to access the data (Tausanovitch, 2020).
 By providing your full name and email address, you will then be emailed a link to the page where it can be downloaded.
@@
 Nationscape samples are provided by Lucid, a market research platform that uses their software for conducting online surveys.
-The Nationscape surveys were completed by participants online in English [tausanovitch_vavreck_reny_hayes_rudkin_2019].
+Nationscape surveys were completed by participants online in English (Tausanovitch, 2020).
 The survey data is then weighted using a simple raking technique, where one set of weights is generated for each week's survey, in an attempt to have data that is representative of their target population.
@@
 typos, misspellings); hence, making it easier to work with.
-This high quality of survey responses can be at least partially attributed to the fact that "Nationscape samples are provided by Lucid, a market research platform that runs an online exchange for survey respondents.
-[@tausanovitch_vavreck_reny_hayes_rudkin_2019]" What is more, the individuals belonging to the sample are required to "complete an attention check before completing the [online] survey [@tausanovitch_vavreck_reny_hayes_rudkin_2019]." With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
+This high quality of survey responses can be at least partially attributed to the fact that "Nationscape samples are provided by Lucid, a market research platform that runs an online exchange for survey respondents(Tausanovitch, 2020)".
+What is more, the individuals belonging to the sample are required to "complete an attention check before completing the [online] survey (Tausanovitch, 2020)." With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
 
@@
 While the goal of the Nationscape survey is to determine American voter attitudes, making inferences based only on the Wave 50 dataset would result in sample bias.
-This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match "a set of demographic criteria [@tausanovitch_vavreck_reny_hayes_rudkin_2019]." It is unlikely that a survey sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
+This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match "a set of demographic criteria (Tausanovitch, 2020)." It is unlikely that a survey sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
 This may lead one to wonder why Nationscape chose to use the described sampling approach.
@@
 Nonetheless, this is a much smaller percentage of non-response in household income than is typically seen in political surveys, and Nationscape handles it by not weighting income for non-respondents.
-Similarly, a statement from Nationscape in its "Representativeness Assessment" implies an overall lack of concern for non-response in the data, as "previous evaluations of the samples Lucid provides have found them to be of high quality [@tausanovitch_vavreck_reny_hayes_rudkin_2019]."
+Similarly, a statement from Nationscape in its "Representativeness Assessment" implies an overall lack of concern for non-response in the data, as "previous evaluations of the samples Lucid provides have found them to be of high quality (Tausanovitch, 2020)."
 
@@
 This allowed for us to look at how age groups might influence the vote.
-The 2020 election marks the first time that Millennials and Gen Z will have an equal share of the number of eligible voters as Baby Boomers and earlier generations [@agearticle] .
+The 2020 election marks the first time that Millennials and Gen Z will have an equal share of the number of eligible voters as Baby Boomers and earlier generations [@agearticle].
 Older voters in the past have had higher turn out rates, so looking at how how age categories tend to vote can give insight on how this will play a factor in the future election.
@@
 # References
-ADD FIVE THIRTY EIGHT CITATION
 
-Tausanovitch, Chris and Lynn Vavreck. 2020. Democracy Fund + UCLA Nationscape, October 10-17, 2019 (version 20200814). Retrieved from [URL].
+Tausanovitch, Chris and Lynn Vavreck. 2020. Democracy Fund + UCLA Nationscape, October 10-17, 2019 (version 20200814). Retrieved from https://www.voterstudygroup.org/publication/nationscape-data-set.
 

### 2020-11-03 outputs/paper/paper.Rmd
bruh

@@
 
-```{r ageind, echo=FALSE, fig.cap="Distribution of Age in Nationscape data"}
+```{r ageind, echo=FALSE, fig.cap="Distribution of Age in Nationscape data", message = FALSE}
 individual_survey <- readRDS(here::here("inputs/cleaned_data","individual-survey.rds"))
@@
 
+s <-cont %>% 
+  select(vote_2016, race) %>% 
+  group_by(race, vote_2016) %>% 
+  summarise(n = n()) %>% 
+  mutate(freq = n/sum(n))
 ```
@@
 #these graph were done as EDA to see if there were any notable trends
-s <-cont %>% 
-  select(vote_2016, race) %>% 
-  group_by(race, vote_2016) %>% 
-  summarise(n = n()) %>% 
-  mutate(freq = n/sum(n))
+
 ## 2016 vote by race
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, include = FALSE}
 # make predictions using our model with data from post-strat data set
@@
 
-## comparison on age
-```{r age, echo=FALSE, include=FALSE, fig.cap="Distribution of Age between ACS and Nationscape data"}
+## Comparison on Age
+```{r age, echo=FALSE, fig.cap="Distribution of Age between ACS and Nationscape data", message = FALSE}
 individual_age_dist<- individual_survey %>% 
@@
 
-## comparison on sex
+## Comparison on Sex
 
-```{r}
+```{r, echo = FALSE,  message = FALSE}
 individual_sex_dist<- individual_survey %>% 
@@
 
-```{r sex_vote,fig.cap="Which individual did the candidate vote for? 2020 on the left, 2016 on the right"}
+```{r sex_vote, fig.cap="Which individual did the candidate vote for? 2020 on the left, 2016 on the right", echo = FALSE, message = FALSE}
 #figure to see 2020 voting by sex
@@
 
-Figure (@fig) indicates that a man is much more likely to vote for Trump than a woman is. This trend is also observed in the 2016 elections when Hillary Clinton, a woman, ran for office.
+(Figure \@ref(fig:sex_vote) indicates that a man is much more likely to vote for Trump than a woman is. This trend is also observed in the 2016 elections when Hillary Clinton, a woman, ran for office.
 
-```{r survey_election_results}
+```{r survey_election_results, echo = FALSE}
 individual_survey %>% mutate(vote_2020 = if_else(vote_2020 == 1,"Donald Trump","Joe Biden")) %>% group_by(vote_2020) %>% tally() %>% rename(candidate = vote_2020, votes = n ) %>% knitr::kable(caption = "Polling results for 2020 Election")
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, message = FALSE}
 

### 2020-11-03 outputs/paper/paper.Rmd
made changes to discussion

--- 2020-11-03 outputs/paper/paper.Rmd
@@
 Our model coefficient results again suggest that men are more likely to vote for Trump, and since we have more women in the ACS data, we would expect a higher percentage of Biden votes.
-This is further supported by other research that suggests that Trump has lost favour with white suburban women, a group he previously was well percevied by[@conroy2020].
+This is further supported by other research that suggests that Trump has lost favour with white suburban women, a group he previously was well perceived by[@conroy2020].
 A lot of models during the 2016 election season suggested a Clinton win, one of the few models that predicted a Trump win was a model that used MRP[@2016election].
@@
 
-Overall, our model predicts a Trump majority. The
+Overall, our model predicts a Trump majority. The model predicts a win in 32 states and a close call in 10 other states. However, it is important to again caution that this is still the popular vote.
 
@@
 
-Weaknesses and next steps should also be included.
+One key weakness in our model is the way we present the results with regards to
 

### 2020-11-03 outputs/paper/paper.Rmd
final commit

@@
 ---
-title: "Prediction of the 2020 Presidential Election"
-subtitle: "Spoiler Alert: Trump Wins"
+title: "Prediction of the 2020 US Presidential Election using MRP on a non-representative poll"
+subtitle: "Predicted Trump win by 52% within a 5% margin of error"
 author: "Anees Shaikh, Jaffa Romain, Lu Mu, and Cameron Fryer"
@@
 The model is then applied to the post-stratification dataset; namely, the results of the 2018 1-year American Community Survey (ACS), to get predictions representative of the target population.
-Since the ACS data pertains to individual persons and their characteristics, the associated use of our model allows us to conclude that Donald Trump will be relected as the President of the United States, winning the popular vote by ___ with a margin of error of ______
-Keywords: forecasting; US 2020 election; Trump; Biden; multilevel regression with post-stratification
+Since the ACS data pertains to individual persons and their characteristics, the associated use of our model allows us to conclude that Donald Trump will be relected as the President of the United States, winning 52% of the popular vote with a margin of error of +-5%.
+  | **Keywords:** forecasting; US 2020 election; Trump; Biden; multilevel regression with post-stratification
+Keywords: "[forecasting; US 2020 election; Trump; Biden; multilevel regression with post-stratification]"
 output:
@@
 
-There are a vast number of ways in which statistical analysis can be sued to foretell future political future  political appointment.
+There are a vast number of ways in which statistical analysis can be used to foretell future political appointment.
 With that being said, however, the statistical methods one chooses to exploit are usually dependent on the data in question.
 As such, it is important to both understand, and be able to describe the data before considering the optimal approach in working with it.
-By staying true to this procedure deriving inferences about some given
+By staying true to this procedure, deriving inferences about some given dataset(s) will be accomplished much more smoothly.
+
+Since the overall goal of this paper is to predict the winner of the 2020 United States Presidential Election using the Democracy Fund + UCLA Nationscape Wave 50 dataset (Tausanovitch and Vavreck, 2020) and the 2018 1-year American Community Survey (ACS) dataset(Ruggles et al, 2020), the first step that we took was becoming familiar with both sets of data.
+Consequently, it became clear that the most effective way to achieve our goal involved employing multi-level regression with post-stratification.
+What brought us to this conclusion is that, although the Nationscape survey contains bias, the survey answers it contained seemed superior to those included in the ACS dataset.
+Furthermore, the sample contained in the ACS data better represented our target population—Americans 18 years and older.
+Thus, we used the Nationscape data to train a model which would then be paired with weightings extracted from the ACS data.
+
+The remainder of the paper is structured as follows: Section 2 provides a summary of the datasets that we used, and highlights their key features; Section 3 involves an explanation of the model we constructed; Section 4  reveals the outcome of our modelling process; and lastly,  Section 5 provides a discussion on our results and notable findings.
+The Github repo is linked in the appendix.
+
 # Data
@@
 The Democracy Fund + UCLA Nationscape is "one of the largest public opinion surveys ever conducted -- interviewing people in nearly every county, congressional district, and mid-sized U.S.
-city in the lead up to the 2020 election (Tausanovitch, 2020)." Consequently, interviewing "roughly 6,250 per week" was part of the survey methodology employed by Nationscape.
+city in the lead up to the 2020 election (Tausanovitch and Vavreck, 2020)." Consequently, interviewing "roughly 6,250 per week" was part of the survey methodology employed by Nationscape.
 As such, the Nationscape Wave 50 dataset contains results for 6,479 interviews conducted during the week of June 25-July 1, 2020.
-It can be accessed by visiting https://www.voterstudygroup.org/publication/nationscape-data-set and submitting a request to access the data (Tausanovitch, 2020).
+It can be accessed by visiting https://www.voterstudygroup.org/publication/nationscape-data-set and submitting a request to access the data (Tausanovitch and Vavreck, 2020).
 By providing your full name and email address, you will then be emailed a link to the page where it can be downloaded.
@@
 Nationscape samples are provided by Lucid, a market research platform that uses their software for conducting online surveys.
-Nationscape surveys were completed by participants online in English (Tausanovitch, 2020).
+Nationscape surveys were completed by participants online in English (Tausanovitch and Vavreck, 2020).
 The survey data is then weighted using a simple raking technique, where one set of weights is generated for each week's survey, in an attempt to have data that is representative of their target population.
@@
 typos, misspellings); hence, making it easier to work with.
-This high quality of survey responses can be at least partially attributed to the fact that "Nationscape samples are provided by Lucid, a market research platform that runs an online exchange for survey respondents(Tausanovitch, 2020)".
-What is more, the individuals belonging to the sample are required to "complete an attention check before completing the [online] survey (Tausanovitch, 2020)." With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
+This high quality of survey responses can be at least partially attributed to the fact that "Nationscape samples are provided by Lucid, a market research platform that runs an online exchange for survey respondents(Tausanovitch and Vavreck, 2020)".
+What is more, the individuals belonging to the sample are required to "complete an attention check before completing the [online] survey (Tausanovitch and Vavreck, 2020)." With this method of sample selection, along with the precaution taken to ensure their attentiveness, it is unlikely for survey respondents to submit low-quality responses.
 
@@
 While the goal of the Nationscape survey is to determine American voter attitudes, making inferences based only on the Wave 50 dataset would result in sample bias.
-This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match "a set of demographic criteria (Tausanovitch, 2020)." It is unlikely that a survey sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
+This is because the target population is all American citizens age 18 years and older, whereas the sample frame is Lucid's online exchange for survey respondents, and the sample is a set of 6,479 individuals from said online exchange which match "a set of demographic criteria (Tausanovitch and Vavreck, 2020)." It is unlikely that a survey sample of 6,479 American citizens during a single week is representative of American voter attitudes as a whole.
 This may lead one to wonder why Nationscape chose to use the described sampling approach.
@@
 Nonetheless, this is a much smaller percentage of non-response in household income than is typically seen in political surveys, and Nationscape handles it by not weighting income for non-respondents.
-Similarly, a statement from Nationscape in its "Representativeness Assessment" implies an overall lack of concern for non-response in the data, as "previous evaluations of the samples Lucid provides have found them to be of high quality (Tausanovitch, 2020)."
+Similarly, a statement from Nationscape in its "Representativeness Assessment" implies an overall lack of concern for non-response in the data, as "previous evaluations of the samples Lucid provides have found them to be of high quality (Tausanovitch and Vavreck, 2020)."
 
@@
 
-```{r model2020prediction}
+```{r model2020prediction,echo=FALSE}
 post_strat_estimates %>% group_by(middle_ground_election_status) %>% tally() %>% rename(state_result = middle_ground_election_status, number = n) %>%  knitr::kable(caption = "Model Prediction for 2020")
@@
 If a candidate wins a state by even a small percent, they effectively win all the electors in the state.
-
-with that background knowledge in mind, what do our poll results indicate?
+With that background knowledge in mind, what do our poll results indicate?
 From Figure(\@ref(fig:survey_election_results)), we gather that Trump has a very slight lead.
@@
 The primary model suggests that Trump has a 91% change of winning, in spite of Trump trailing Biden in the polls[@norpoth_2020].
-
-Overall, our model predicts a Trump majority. The model predicts a win in 32 states and a close call in 10 other states. However, it is important to again caution that this is still the popular vote.
+Overall, our model predicts a Trump majority.
+The model predicts a win in 32 states and a close call in 10 other states.
+However, it is important to again caution that this is still the popular vote.
 
@@
 
-One key weakness in our model is the way we present the results with regards to
+One key weakness in our model is the way we present the results with regards to states won.
+If we look at it from a conservative estimate point of view, where we only use the lower bounds, then we'll see a vastly different election result.
+We chose to remove state from the model as nearly no state had a significant value in its coeffiecents.
+This model could be enhanced by using a different poll.
+For instance @2016_election uses an A+ quality survey, as determined by FiveThirtyEight, and was able to best predict the 2016 election.
+We could try that approach and see if the result changes.
+The number of dimensions we have are too few, there were a few variables within the Nationscape dataset such as voting behaviour in 2016 that we weren't able to get a similar value for from the post-stratified dataset.
+Such variables likely have high explanatory power and would make our model overall much more effective.
+A 5% margin of error is quite large, especially when elections are extremely close.
+A reduction in the margin of error would result in being able to much more accurately predict the election.
 
@@
 
-\newpage
+All code can be found at: https://github.com/aneesshake/elections_prediction
 
